---
title: "Pepe Annual Data Analysis"
author: "Michelle DePrenger-Levin"
date: "2024-06-26"
output: R script
---


2024:    
      Transect 15: tag lost so new origin tag put in. Transect 15 starts new in 2024    
      Transect 16 lost completely. No data for Transect 16        
      Transect 21 is lost     
      
2025:    
      Found some plots, lost another?? Lost one of them

```{r}

# rm(list=ls())
library(R2jags)
library(runjags)
# library(mcmcplots)
library(boot)
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(RMark)

library(lme4)
require(AICcmodavg)

library(myClim)

setwd("C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii")

```


Import data     

Move to GitHub once checked and correct  
```{r}

trans2024 <- read.csv("Data/20240711_dbExport/20240711__pepe_triState_transects.csv", header = TRUE)

## Once this gets corrected, for now use old
plants2024 <- read.csv("Data/20240712_dbExport/20240712__pepe_triState_plants.csv", header = TRUE)

## Need to check datasheets, then remove plantID 139
plants2024 <- plants2024 %>%
  filter(plantID != 139)

data2024 <- read.csv("Data/20240712_dbExport/20240712_pepe_triState_data.csv")

surv2023_2024 <- data2024 %>%
  left_join(plants2024, by = "plantID")


head(surv2023_2024)

### Add 2025
load("Data/2025_NewPlants.Rdata")
# loads as "allDataFix" from Data_qualitycontrol.Rmd 
plants2025 <- allDataFix

## The reproductive status is "1" if there are any fruit
plants2025 <- plants2025 %>%
  mutate(Reproductive.Status = case_when(Number.Reproductive.Structures > 0 ~ 1,
                                         TRUE ~ Reproductive.Status))

## Merge with transplant and transplant season from plants2024
plants2025 <- plants2025 %>%
  # dplyr::select(c(plantID, date, transectID, x, y, Height.in.cm:recordedBy, transplantSeason, transectPosition, isTransplant)) %>%
  mutate(transectID =as.integer(transectID)) %>%
  replace_na(list(reproductiveStatus = 0, numberReproductiveStructures = 0))

plants2025 <- plants2025 %>%
  left_join(plants2024) %>%
  left_join(trans2024[,-1], by = c("transectID" = "transectId")) 

## Clean up and standardize NULL and NA
plants2025 <- plants2025 %>%
  mutate(transplantSeason = case_when( !(transplantSeason %in% c("Spring", "Fall")) ~ NA,
                                       TRUE ~ transplantSeason)) %>%
  mutate(transectPosition = as.numeric(transectPosition)) %>%
  replace_na(list(isTransplant = 0))

```


# Check correct assigment of PlantIDs to transplants (and subsequently all of them)   
```{r}
## Five Fall transplants, one transect, was measured during tranplanting and during plot reading. 
table(surv2023_2024$date, surv2023_2024$transplantSeason)

### Change all dates to one in spring, one in fall and one in 2024 
surv2023_2024 <- surv2023_2024 %>%
  mutate(date = case_when(date == "2023-10-17" ~ "2023-10-16",
                          date == "2024-06-25" ~ "2024-06-24",
                          TRUE ~ date))
## Now looks good, Except are there duplicate fall records? and some additional records for both fall and spring in 2024??
table(surv2023_2024$date, surv2023_2024$transplantSeason)

## I did double observations in 2024. For now keep one, didn't get double in 2024, or do first and second measurement
surv2023_2024 <- surv2023_2024 %>%
  mutate(recordedBy = case_when(recordedBy == "Michael Bone, Stanton Schell" ~ "MB_SS",
                                recordedBy == "Mike Bone, Stanton Schell" ~ "MB_SS",
                                recordedBy == "Arich Fruehling, Leah Veldhuisen" ~ "AF_LV",
                                recordedBy == "Geena Poulter, Michelle DePrenger-Levin" ~ "GP_MEDL",
                                recordedBy == "Michelle DePrenger-Levin" ~ "MEDL",
                                recordedBy == "Michelle DePrenger-Levin, Adriana Jacobi" ~ "MEDL_AJ",
                                recordedBy == "Michael Guidi, Brooke Washburn, Syed Jalalzai" ~ "MG_BW_SJ"
                                ))

surv2023_2024 %>%
  filter(transplantSeason == "Fall") %>%
  group_by(date, transectID, recordedBy) %>%
  add_count() %>%
  dplyr::select(plantID:heightInCentimeters, x,y ,recordedBy , transectID, transplantSeason:isTransplant, n) %>%
  arrange(transectID, transectPosition) 

surv2023_2024 %>%
  mutate(transectPosition = as.numeric(transectPosition)) %>%
  filter(date == "2023-10-16" & transectID == 34) %>%
  filter(isTransplant == 1) %>%
  arrange(transectPosition) %>%
  group_by(transplantSeason, transectPosition) %>%
  add_count() %>%
  dplyr::select(plantID:heightInCentimeters, transectID:n)


## Two plants measured twice
surv2023_2024 %>%
  filter(date == "2023-05-31") %>%
  group_by(transectID, recordedBy, isTransplant, plantID) %>%
  add_count() %>%
  filter(n > 1)

surv2023_2024 <- surv2023_2024 %>%
  filter(!(id.x %in% c(187,189)))


surv2023_2024 <- surv2023_2024 %>%
  dplyr::select(plantID:recordedBy, transectID:isTransplant)

```

Missing 16 and 21 in 2024
```{r}
surv2023_2024 %>%
  filter(transectID == 16 & date == "2024-06-24")

surv2023_2024 %>%
  filter(transectID == 21 & date == "2024-06-24")

## No 15, no 21
plants2025 %>%
  filter(date != "plant not recorded in 2025") %>%
  group_by(transectID) %>%
  summarise(n = n()) %>%
  print(n = 40)

# ## Remove Transect 16
# surv2023_2024 <- surv2023_2024 %>%
#   filter(transectID != 16)

## Need to add a 1st and 2nd observation for 2024. Second for only some plots in 2023, some weren't measured (I think < transect 10??)
## Yes, transects 1-9 we only measured plants that were the transplants in the spring. 
surv2023_2024 %>%
  filter(date == "2023-05-31") %>%
  group_by(transectID, recordedBy, isTransplant) %>%
  summarise(n = n()) %>%
  arrange(transectID) %>% 
  filter(isTransplant == 0) %>%
  print(n = 32)


## Maybe we successfully found small plots with no plants for the treatments? at least under 10
# nothing in 2,4,6,8,14,28,30
surv2023_2024 %>%
  filter(date == "2023-10-16") %>%
  group_by(transectID, recordedBy, isTransplant) %>%
  summarise(n = n()) %>%
  arrange(transectID) %>% 
  filter(isTransplant == 0) %>%
  print(n = 33)


# Even are transplants, Odds are controls
# No existing in 6, 15 lost, 21 lost, nothing in 28, Nothing in 30. 
plants2025 %>%
  filter(date != "plant not recorded in 2025") %>%
  group_by(transectID, recordedBy, isTransplant) %>%
  summarise(n = n()) %>%
  arrange(transectID) %>% 
  filter(isTransplant == 0) %>%
  print(n = 39)

# Remove one Transect 18, no recorder?
# ?? transect 19 28 with no recorder, only 1 under BW and MG, assign BW; MG to 19
# 20 should all be MG; BW
## All the missing observers are fine, they just did something funny entering data. 
plants2025 %>%
  filter(date != "plant not recorded in 2025") %>%
  group_by(transectID, recordedBy) %>%
  summarise(n = n()) %>%
  arrange(transectID) %>% 
  print(n = 49)


plants2025 %>%
  filter(date != "plant not recorded in 2025") %>%
  filter(recordedBy == "" & Height.in.cm == "")

```

Measurements:
        1st                                                     2nd              
2023:  10,12,16,18,20,22,24,26,34,36,38,40               all other than  2,4,6,8,14,28,30     
2024:  all but 16 and 21                                 all but 16 and 21
2025: none have double    

## First time only one observation per year; average measurements between two groups, or keep max if only one team observed. 
## OR two times in 2024, once all other years??
```{r}

plants2025 %>%
  filter(date != "plant not recorded in 2025") %>%
  group_by(transectID, recordedBy) %>%
  summarise(n = n()) %>%
  print(n=47)

plants2025 %>%
  filter(recordedBy == "" & Height.in.cm > 0) %>%
  group_by(transectID) %>%
  summarise(n=n())

### Fill recordedBy for those with missing

plants25toadd <- plants2025 %>%
  filter(recordedBy != "") %>%
  group_by(transectID) %>%
  mutate(recordedBy = unique(recordedBy)[1]) %>%
  ungroup() %>%
  dplyr::select(plantID,date, Height.in.cm:Comments, recordedBy, transectID, x, y, transplantSeason:isTransplant) %>%
  rename(heightInCentimeters = Height.in.cm) %>%
  rename(basalWidthInCentimeters = Basal.Width.in.cm) %>%
  rename(numberReproductiveStructures = Number.Reproductive.Structures) %>%
  rename(reproductiveStatus = Reproductive.Status) %>%
  rename(comments = Comments) %>%
  filter(date != "plant not recorded in 2025") %>%
  mutate(recordedBy = case_when(recordedBy == "" ~ NA,
                                TRUE ~ recordedBy)) %>%
  arrange(transectID, recordedBy) %>%
  fill(recordedBy) 

plants25toadd %>%
  group_by(transectID, recordedBy) %>%
  summarise(n = n()) %>%
  print(n = 47)

surv_wide <- 
  surv2023_2024 %>%
  mutate(transplantSeason = case_when( !(transplantSeason %in% c("Spring", "Fall")) ~ NA,
                                       TRUE ~ transplantSeason)) %>%
  mutate(transectPosition = as.numeric(transectPosition)) %>%
  bind_rows(plants25toadd) %>%
    mutate(date = case_when(date == "6/25/2025" ~ "6/24/2025",
                            ## To make averages for the few doubles in 2023
                            date == "2023-05-31" ~ "2023-10-16",
                            TRUE ~ date)) %>%
  mutate(Date = as.Date(date)) %>%
  filter(!(transectID %in% c(15, 21,16))) %>%
  ## Can ignore the 5 repeat collections of Fall transplants
  filter(recordedBy != "MEDL_AJ") %>%
  # filter(!(transectID %in% c(15,16, 21))) %>%
  select(date,plantID:numberReproductiveStructures,transectID:isTransplant) %>%
  arrange(date) %>%
  group_by(plantID, date) %>%
  add_count() %>%
    mutate(heightInCentimeters = case_when(n == 2 ~ mean(heightInCentimeters),
                                           n == 1 ~ max(heightInCentimeters)),
           basalWidthInCentimeters = case_when(n == 2 ~ mean(basalWidthInCentimeters),
                                           n == 1 ~ max(basalWidthInCentimeters)),
           reproductiveStatus = case_when(n == 2 ~ max(basalWidthInCentimeters),
                                           n == 1 ~ max(basalWidthInCentimeters)),
           numberReproductiveStructures = case_when(n == 2 ~ mean(numberReproductiveStructures),
                                           n == 1 ~ max(numberReproductiveStructures))) %>%
  ungroup() %>%
    distinct() %>%
  mutate(ch = case_when(heightInCentimeters > 0 ~ 1,
                        heightInCentimeters == 0 ~ 0)) %>%
  pivot_wider(names_from = date,
              id_cols = c(plantID, transectID, transplantSeason, isTransplant),
              values_from = ch,
              # values_from = Obs,
              # values_fill = "0"
              ) %>%
  arrange(transectID,transplantSeason) %>%
  distinct()


## CJS
surv_wide <- surv_wide %>%
  replace_na(list(`2023-10-16` = 0,
                  `2024-06-24` = 0,
                  `6/24/2025` = 0)) %>%
  unite(ch, `2023-10-16`:`6/24/2025`, sep = "")


## Remove the four without alive ever
surv_wide %>%
  filter(!grepl("1",ch))

surv_wide <- surv_wide %>%
  filter(grepl("1", ch))


```


## Shouldn't penalize plots for having transplants with lower survival 
```{r}
pepe_ch <- surv_wide %>%
  filter(isTransplant != 1)%>%
              left_join(trans2024[,2:4], by = c("transectID" = "transectId")) %>%
  mutate(Trans = as.factor(transectID),
         Type = as.factor(type_CT_TT)) %>%
  filter(!is.na(Trans))

pepe_ch %>%
  filter(isTransplant != 1)

plotdf <- pepe_ch %>%
  distinct(Trans) 

dmPlot <-model.matrix(~ -1 + Trans, pepe_ch)
dimnames(dmPlot)[[2]][length(dimnames(dmPlot)[[2]])] ## The last one to add a semicolon to

pepe_ch %>%
  dplyr::select(ch) %>%
  bind_cols(dmPlot) %>%
  mutate(Trans40 = paste(Trans40, ";", sep= "")) %>%
  write.table(file = "Data/Pepe_mark.inp", sep = " ", col.names = FALSE, row.names = FALSE)

pepeCJS <- convert.inp("Data/Pepe_mark.inp",
                       group.df = plotdf,
                       covariates = NULL,
                       use.comments = FALSE)




```


Cormick-Jolly-Seber
```{r}

pepe.pr <- process.data(pepeCJS, model = "CJS", begin.time = 2023, groups = "Trans")
pepe.dd <- make.design.data(pepe.pr)

## Need to index the plots not seen and fix them to 0. Transects 15 (starts new 2024), 16, and 21, just 16 in 2025, remove until 2026 data
# pindex <- pepe.dd$p
# p.missing16 <- pindex %>%
#   mutate(cohort = as.numeric(as.character(cohort))) %>%
#   filter(group %in% c(16) & cohort == 2024)

## Fixed indices
# p.indices <- c(p.missing16$par.index)
# p.values <- rep(0, length(pindex))


CJS.pepe <- mark(pepe.pr, pepe.dd, model = "CJS",
                 model.parameters = list(p = list(formula = ~ 1),
                                                    # fixed = list(index = p.indices, value = p.values),
                                         Phi = list(formula = ~ 1 + Trans)))


TransTreatdf <- pepe_ch %>%
  distinct(Trans, Type, block) 

Phi.table <- get.real(CJS.pepe, "Phi", se = TRUE) %>%
  distinct(estimate, lcl, ucl, .keep_all = TRUE) %>%
  left_join(TransTreatdf) 

## Difference in survival between Control and Transplant, color is by block
ggsave(filename = "Figures/SurvivalCJS_block.jpg",

ggplot(Phi.table, aes(Type, estimate, color = as.factor(block))) +
  geom_point(position = position_dodge(width = .3)) +
  geom_errorbar(aes(ymin = lcl, ymax = ucl),position = position_dodge(width = .3))+
  theme_bw() +
  ylab("Survival")+
  xlab("Transect") +
  guides(color = "none")+
  scale_x_discrete(labels = c("Control","Transplant"))


, width=120, height=100,units='mm', dpi=300)  

Phi.table

Phi.table %>%
  group_by(Type) %>%
  summarise(est = mean(estimate),
            lcl = mean(lcl),
            ucl = mean(ucl))
```


## Transplants compared to in situ  
```{r}

surv_wide %>%
  replace_na(list(isTransplant = 0)) %>%
  group_by(isTransplant) %>%
  summarise(n = n())

pepe_ch_transplant <- surv_wide %>%
  replace_na(list(isTransplant = 0)) %>%
              left_join(trans2024[,2:4], by = c("transectID" = "transectId")) %>%
  mutate(Trans = as.factor(transectID),
         Type = as.factor(type_CT_TT),
         Transplant = as.factor(isTransplant)) %>%
  replace_na(list(isTransplant = 0)) 

transplantdf <- pepe_ch_transplant %>%
  distinct(Transplant) 

dmTransplant <-model.matrix(~ -1 + Transplant, pepe_ch_transplant)
dimnames(dmTransplant)[[2]][length(dimnames(dmTransplant)[[2]])] ## The last one to add a semicolon to

pepe_ch_transplant %>%
  dplyr::select(ch) %>%
  bind_cols(dmTransplant) %>%
  mutate(Transplant1 = paste(Transplant1, ";", sep= "")) %>%
  write.table(file = "Data/Pepe_Transplant_mark.inp", sep = " ", col.names = FALSE, row.names = FALSE)

pepeCJS <- convert.inp("Data/Pepe_Transplant_mark.inp",
                       group.df = transplantdf,
                       covariates = NULL,
                       use.comments = FALSE)




```


Cormack-Jolly-Seber
```{r}

pepe.pr <- process.data(pepeCJS, model = "CJS", begin.time = 2023, groups = "Transplant")
pepe.dd <- make.design.data(pepe.pr)

## Need to index the plots not seen and fix them to 0. Transects 15 (starts new 2024), 16, and 21, just 16 in 202, remove until 2026 data
# pindex <- pepe.dd$p
# p.missing16 <- pindex %>%
#   mutate(cohort = as.numeric(as.character(cohort))) %>%
#   filter(group %in% c(16) & cohort == 2024)

## Fixed indices
# p.indices <- c(p.missing16$par.index)
# p.values <- rep(0, length(pindex))


CJS.pepe <- mark(pepe.pr, pepe.dd, model = "CJS",
                 model.parameters = list(p = list(formula = ~ 1),
                                                    # fixed = list(index = p.indices, value = p.values)),
                                         Phi = list(formula = ~ 1 + Transplant)))

Phi.table <- get.real(CJS.pepe, "Phi", se = TRUE) %>%
  distinct(estimate, lcl, ucl, .keep_all = TRUE) 
## Difference in survival between Control and Transplant, color is by block
ggsave(filename = "Figures/SurvivalCJS_TransplantInSitu.jpg",

ggplot(Phi.table, aes(Transplant, estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = lcl, ymax = ucl))+
  theme_bw() +
  ylab("Survival")+
  xlab("") +
  guides(color = "none")+
  scale_x_discrete(labels = c("In-situ","Transplant"))


, width=120, height=100,units='mm', dpi=300)  

# The few transplanted individuals compared to all the in-situ plants. 
Phi.table
```


################################################################################################################

## Among transplants, fall vs. spring
```{r}
## Need to remove the 4? repeats? Maybe some of the transplants that aren't really got added? 
surv_wide %>%
  replace_na(list(isTransplant = 0,
                  transplantSeason = "NULL")) %>%
  group_by(isTransplant, transplantSeason) %>%
  summarise(n = n())

## Missing 5 individuals (Oh! the lost transect!!) 
surv_wide %>%
  group_by(isTransplant, transplantSeason) %>%
  summarise(n = n())

pepe_ch_trSeason <- surv_wide %>%
  replace_na(list(isTransplant = 0)) %>%
              left_join(trans2024[,2:4], by = c("transectID" = "transectId")) %>%
  filter(isTransplant == 1) %>%
  mutate(Trans = as.factor(transectID),
         Type = as.factor(type_CT_TT),
         TransplantSeason = as.factor(transplantSeason)) %>%
  replace_na(list(isTransplant = 0))           

transSeasontdf <- pepe_ch_trSeason %>%
  distinct(TransplantSeason) 

dmTransSeason <-model.matrix(~ -1 + TransplantSeason, pepe_ch_trSeason)
dimnames(dmTransSeason)[[2]][length(dimnames(dmTransSeason)[[2]])] ## TransplantSeasonSpring

pepe_ch_trSeason %>%
  dplyr::select(ch) %>%
  bind_cols(dmTransSeason) %>%
  mutate(TransplantSeasonSpring = paste(TransplantSeasonSpring, ";", sep= "")) %>%
  write.table(file = "Data/Pepe_TransplantSeason_mark.inp", sep = " ", col.names = FALSE, row.names = FALSE)

pepeCJS <- convert.inp("Data/Pepe_TransplantSeason_mark.inp",
                       group.df = transSeasontdf,
                       covariates = NULL,
                       use.comments = FALSE)



pepe.pr <- process.data(pepeCJS, model = "CJS", begin.time = 2023, groups = "TransplantSeason")
pepe.dd <- make.design.data(pepe.pr)

## Need to index the plots not seen and fix them to 0. Transects 15 (starts new 2024), 16, and 21, just 16 in 202, remove until 2026 data
# pindex <- pepe.dd$p
# p.missing16 <- pindex %>%
#   mutate(cohort = as.numeric(as.character(cohort))) %>%
#   filter(group %in% c(16) & cohort == 2024)

## Fixed indices
# p.indices <- c(p.missing16$par.index)
# p.values <- rep(0, length(pindex))


CJS.pepe <- mark(pepe.pr, pepe.dd, model = "CJS",
                 model.parameters = list(p = list(formula = ~ 1),
                                                    # fixed = list(index = p.indices, value = p.values)),
                                         Phi = list(formula = ~ 1 + TransplantSeason)))

Phi.table <- get.real(CJS.pepe, "Phi", se = TRUE) %>%
  distinct(estimate, lcl, ucl, .keep_all = TRUE) 
## Difference in survival between Control and Transplant, color is by block
ggsave(filename = "Figures/SurvivalCJS_TransplantSeason.jpg",

ggplot(Phi.table, aes(TransplantSeason, estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = lcl, ymax = ucl))+
  theme_bw() +
  ylab("Survival")+
  xlab("") +
  guides(color = "none")


, width=120, height=100,units='mm', dpi=300)

# The few transplanted individuals compared to all the in-situ plants. 
Phi.table
```

####################################################################################################################



Robust design and multistate   
```{r}

##Need doubles in 2023 as well for the ones transect 10 on, that got measured

survey2bind <- surv2023_2024 %>%
  mutate(transplantSeason = case_when( !(transplantSeason %in% c("Spring", "Fall")) ~ NA,
                                       TRUE ~ transplantSeason)) %>%
  mutate(transectPosition = as.numeric(transectPosition)) %>%
  bind_rows(plants25toadd) %>%
  filter(transectID != "") %>%
    mutate(date = case_when(date == "6/25/2025" ~ "2025-06-24",
                            date == "6/24/2025" ~ "2025-06-24",
                            TRUE ~ date)) %>%
  mutate(Date = as.Date(date)) %>%
  mutate(Year = format(as.Date(Date, format="%Y-%m-%d"),"%Y")) %>%
  filter(!(transectID %in% c(15, 21,16))) %>%
  ## Can ignore the 5 repeat collections of Fall transplants
  filter(recordedBy != "MEDL_AJ") %>%
  left_join(trans2024, by = c("transectID" = "transectId")) %>%
  select(Date,Year,plantID:numberReproductiveStructures,transectID:transplantSeason,isTransplant, recordedBy) %>%
  mutate(surveyID = as.numeric(as.factor(paste(date,recordedBy)))) %>%
  arrange(Date) %>%
  group_by(transectID,surveyID,Date,Year,recordedBy) %>%
  summarise(plants = n()) %>%
  group_by(transectID, Year) %>%
  mutate(survID = row_number()) %>%
  mutate(surveyID = case_when(Date == "2023-05-31" ~ 1,
                              Date == "2023-10-16" ~ 2,
                              (survID == 1 & Date == "2024-06-24") ~ 3, ## and the other one is four
                              (survID == 2 & Date == "2024-06-24") ~ 4, ## and the other one is four
                              Date == "2025-06-24" ~ 5,
                              TRUE ~ surveyID))


survey2bind %>% print(n = 10)
  
## Now add number for 2023 survey 1, 2024 has surveys 2 and 3 when doubled, 2025 has survey 4
surv_wide_robust <-
  surv2023_2024 %>%
  mutate(transplantSeason = case_when( !(transplantSeason %in% c("Spring", "Fall")) ~ NA,
                                       TRUE ~ transplantSeason)) %>%
  mutate(transectPosition = as.numeric(transectPosition)) %>%
  bind_rows(plants25toadd) %>% 
  filter(transectID != "") %>%
    mutate(date = case_when(date == "6/25/2025" ~ "2025-06-24",
                            date == "6/24/2025" ~ "2025-06-24",
                            TRUE ~ date)) %>%
  mutate(Date = as.Date(date)) %>%  # group_by(Date) %>% summarise(n = n()) 
  mutate(Year = format(as.Date(Date, format="%Y-%m-%d"),"%Y")) %>%
  filter(!(transectID %in% c(15, 21,16))) %>%
  ## Can ignore the 5 repeat collections of Fall transplants
  filter(recordedBy != "MEDL_AJ") %>%
  left_join(trans2024, by = c("transectID" = "transectId")) %>%
  select(Date,plantID:numberReproductiveStructures,transectID:transplantSeason,isTransplant, recordedBy, type_CT_TT) %>%
  left_join(data.frame(survey2bind[,c(1:5)]), by = c("transectID" = "transectID", "recordedBy" = "recordedBy", 
                                                     "Date" = "Date")) %>%
    filter(heightInCentimeters > 0 & !is.na(heightInCentimeters)) %>%
    # group_by(surveyID, transectID, type_CT_TT) %>%
    # summarise(n = n()) %>%
    # ggplot(  aes(surveyID, n, color = as.factor(transectID)))+
    # geom_point(position = position_dodge(width = 0.1)) +
    # geom_line(position = position_dodge(width = 0.1)) +
    # facet_wrap(~ type_CT_TT)
  ### Going to need to set p to zero for the ones with only one measurement in 2023 and 2024
  arrange(surveyID) %>%
  replace_na(list(reproductiveStatus = 0)) %>%
  mutate(Obs = case_when( (heightInCentimeters > 0 & reproductiveStatus == 0) ~ "V",
                          (heightInCentimeters > 0 & reproductiveStatus == 1) ~ "R",
                          heightInCentimeters == 0 ~ "0"
                          # ,
                          # is.na(heightInCentimeters) ~ "0"
                          )) %>%
  # dplyr::select(plantID,surveyID,Obs, transectID, transplantSeason, isTransplant, recordedBy) %>% 
    ## One accidental duplicate record in the same survey, same observers
    # group_by(plantID, surveyID) %>%
    # add_count() %>%
    # filter(n > 1) 
    filter(!(plantID == 1629 & surveyID == 3 & Obs == "0")) %>%
    pivot_wider(names_from = surveyID,
              id_cols = c(plantID,transectID,isTransplant,transplantSeason), ## The recorded by is only each survey
              values_from = Obs,
              values_fill = "0",
              names_prefix = "Survey"
              ) %>%
  arrange(transectID, plantID, transplantSeason) %>%
  unite(ch, Survey1:Survey5, sep = "")

surv_wide_robust %>% data.frame() %>% head(20)
plants2025 %>% filter(plantID == 2057) ## Fixed by adding replace_na fro reproductiveStatus

## Any that are always dead like AsMi?? Nope, all look good, have at least one V or R
surv_wide_robust %>%
  filter(!(grepl("R",ch) | grepl("V",ch)))

surv_wide_robust %>% head()

```


Need to fix indeces for those that don't have double in 2024, only some have double, same for 2023  
```{r}
surv2023_2024 %>%
  group_by(transectID, date, recordedBy) %>%
  summarise(n = n()) %>%
  # group_by(transectID) %>%
  # summarise(surveys = n()) %>%
  # filter(surveys == 4) %>%
  print(n = 150)
```

Transect    surveys
1             2,3,4,5            
2           1,2,3,4,5           
3             2,3,4,5            
4           1,2,3,4,5            
5             2,3,4,5               
6           1,2,3,4,5            
7             2,3,4,5               
8           1,2,3,4,5         
9             2,3,4,5               
10          1,2,3,4,5                
11            2,3,4,5               
12          1,2,3,  5              
14          1,2,3,4,5         
15            2,3,4     New nail in 3 and 4, Then lost again!!!, remove
16          1,2,    5   No data in 2024
17            2,3,4,5         
18          1,2,3,4,5         
19            2,3,4,5         
20          1,2,3,  5         
21                      lost, remove
22          1,2,3,4,5         
23            2,3,4,5         
24          1,2,3,4,5         
25            2,3,  5         
26          1,2,3,4,5         
27            2,3,4,5         
28          1,2,3,4,5         
29            2,3,4,5         
30          1,2,3,4,5         
31            2,3,  5         
32          1,2,3,  5         
34          1,2,3,4,5         
35            2,3,  5         
36          1,2,3,  5         
37            2,3,4,5         
38          1,2,3,  5         
39            2,3,  5         
40          1,2,3,  5   


```{r}

## Three have NA pepe_multi[1038,], 1075, and 1387
# pepe_multi[c(1038,1075,1387),]
## Plants with NA at the third (1165 and 732) and second 928 survey
finderrors <- c(1165,732,928)
surv2023_2024 %>%
  filter(plantID %in% finderrors) %>%
  mutate(date = as.Date(date)) %>%
  arrange(plantID, date)

surv_wide_robust %>%
  filter(plantID %in% finderrors)

pepe_multi <- surv_wide_robust %>%
  filter(!(plantID %in% finderrors)) %>%
  left_join(trans2024[,c(2:4)], by = c("transectID" = "transectId")) %>%
  filter(isTransplant != 1) %>%
  mutate(Trans = as.factor(transectID),
         Type = as.factor(type_CT_TT)) %>%
  filter(!is.na(Trans))

pepe_multi %>%
  filter((grepl("R",ch) | grepl("V",ch)))
  

dmPlot <-model.matrix(~ -1 + Trans, pepe_multi)
dimnames(dmPlot)[[2]][length(dimnames(dmPlot)[[2]])] ## The last one to add a semicolon to

pepe_multi %>%
  dplyr::select(ch) %>%
  bind_cols(dmPlot) %>%
  mutate(Trans40 = paste(Trans40, ";", sep= "")) %>%
  write.table(file = "Data/Pepe_mark_multistate.inp", sep = " ", col.names = FALSE, row.names = FALSE)

pepeCJS <- convert.inp("Data/Pepe_mark_multistate.inp",
                       group.df = plotdf,
                       covariates = NULL,
                       use.comments = FALSE)  


pepe.pr <- process.data(pepe_multi, model = "Multistrata", begin.time = 2023, groups = "Trans",
                               strata.labels = c("R","V"))
pepe.dd <- make.design.data(pepe.pr)

## Need to index the plots not seen and fix them to 0. Transects 15 (starts new 2024), 16, and 21, just 16 in 202, remove until 2026 data
pindex <- pepe.dd$p
p.missing16 <- pindex %>%
  mutate(cohort = as.numeric(as.character(cohort))) %>%
  filter(group %in% c(16) & cohort == 2024)

## Fixed indices
# p.indices <- c(p.missing16$par.index)
# p.values <- rep(0, length(pindex))


CJS.pepe <- mark(pepe.pr, pepe.dd, model = "CJS",
                 model.parameters = list(p = list(formula = ~ 1),
                                                    # fixed = list(index = p.indices, value = p.values)),
                                         Phi = list(formula = ~ 1 + Trans)))



```


## Something is wrong
```{r}

plants25toadd <- plants25toadd %>%
  mutate(year = format(as.Date(date, format = "%m/%d/%Y"), "%Y")) 
  

######################### Appendices ###############################
surv2023_2024 %>%
  mutate(Date = as.Date(date)) %>%
  ## Multiple measurements per year
  mutate(year = format(Date, "%Y")) %>%
  bind_rows(plants25toadd) %>%
  filter(isTransplant == 0) %>% 
  left_join(trans2024, by = c("transectID" = "transectId")) %>%
  filter(!(transectID %in% c(15, 16, 21))) %>%
  arrange(Date) %>%
  group_by(year, transectID, type_CT_TT, plantID, x, y) %>%
  summarise(heightInCentimeters = max(heightInCentimeters, na.rm = TRUE),
            basalWidthInCentimeters = max(basalWidthInCentimeters, na.rm = TRUE),
            reproductiveStatus = max(reproductiveStatus, na.rm = TRUE),
            numberReproductiveStructures = max(numberReproductiveStructures, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(surv = case_when(heightInCentimeters > 0 ~ 1,
                          heightInCentimeters == 0 ~ 0)) %>%
  ## All have to be alive in the first year
  filter(!(year == 2023 & surv == 0)) %>%
  write.csv("2023_2025_Penstemon-penlandii_existing.csv")

surv2023_2024 %>%
  filter(isTransplant == 1) %>% 
  left_join(trans2024, by = c("transectID" = "transectId")) %>%
  # filter(!(transectID %in% c(15, 16, 21))) %>%
  mutate(Date = as.Date(date)) %>%
  arrange(Date) %>%
  ## Multiple measurements per year
  mutate(year = format(Date, "%Y")) %>%
  group_by(year, transectID, type_CT_TT, plantID, x, y, transplantSeason, transectPosition) %>%
  summarise(heightInCentimeters = max(heightInCentimeters, na.rm = TRUE),
            basalWidthInCentimeters = max(basalWidthInCentimeters, na.rm = TRUE),
            reproductiveStatus = max(reproductiveStatus, na.rm = TRUE),
            numberReproductiveStructures = max(numberReproductiveStructures, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(surv = case_when(heightInCentimeters > 0 ~ 1,
                          heightInCentimeters == 0 ~ 0)) %>%
  ## All have to be alive in the first year
  filter(!(year == 2023 & surv == 0)) %>%
  write.csv("C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026//2024_Penstemon-penlandii_Annualreport/2023_2024_Penstemon-penlandii_transplants.csv")

#############################################################################################


```


Percent of individuals that survived by 2025 per transect and reproductive status by transplant timing  
```{r}

ggplot()


```









































Pre 2025


# Survival   
```{r}

# wide format for detection
surv_wide <- surv2023_2024 %>%
  mutate(Date = as.Date(date)) %>%
  filter(!(transectID %in% c(15,16, 21))) %>%
  arrange(Date) %>%
  ## Can ignore the 5 repeat collections of Fall transplants
  filter(recordedBy != "Michelle DePrenger-Levin, Adriana Jacobi") %>%
  mutate(ch = case_when(heightInCentimeters > 0 ~ 1,
                        heightInCentimeters == 0 ~ 0)) %>%
  mutate(Survey = case_when(date == "2023-05-31" ~ "aSp23",
                            date == "2023-10-16" ~ "bFa23",
                            date == "2024-06-24" ~ "cSm24")) %>%
  mutate(Obsr = case_when(recordedBy == "Michelle DePrenger-Levin" ~ "MEDL",
                          # recordedBy == "Michelle DePrenger-Levin, Adriana Jacobi" ~ "MEDL_AJ",
                          recordedBy == "Mike Bone, Stanton Schell" ~ "MB_SS",
                          recordedBy == "Arich Fruehling, Leah Veldhuisen" ~ "AF_LV",
                          recordedBy == "Geena Poulter, Michelle DePrenger-Levin" ~ "GP_MEDL",
                          recordedBy == "Michael Guidi, Brooke Washburn, Syed Jalalzai" ~ "MG_BW_SJ")) %>%
  pivot_wider(names_from = c(Survey, Obsr),
              values_from = ch,
              id_cols = c(plantID,transplantSeason,transectID) 
               ,values_fn = function(x) paste(x, collapse = ", ")# first  ## Because why? why are there double for the same date and observers?
              ) %>%
  arrange(transectID,transplantSeason)



### Combine columns across surveys ignoring the NAs  
## Need to add zeros to all the fall transplants during the spring which I think I do with the mutate

surv_wide %>%
  mutate(across(aSp23_MEDL:bFa23_MEDL, replace_na, "0")) %>%
  unite(ch23, aSp23_MEDL:bFa23_MEDL , sep = "") %>% 
  unite(ch24, cSm24_MB_SS:cSm24_MG_BW_SJ, sep = "") %>%
  unite(ch, ch23:ch24, sep = "") %>%
  filter(nchar(ch) < 11) %>%
  left_join(plants2024, by = "plantID") %>%
  left_join(data2024, by = "plantID")
  
pepe_ch <- surv_wide %>%
  mutate(across(aSp23_MEDL:bFa23_MEDL, replace_na, "0")) %>%
  unite(ch23, aSp23_MEDL:bFa23_MEDL , sep = "") %>% 
  unite(ch24, cSm24_MB_SS:cSm24_MG_BW_SJ, sep = "", na.rm = TRUE) %>%
  unite(ch, ch23:ch24, sep = "")

## Only two characters would be not measured at all in 2024; still 1753 missing 
pepe_ch %>%
  filter(nchar(ch) < 3) %>%
  left_join(plants2024, by = "plantID") %>%
  filter(isTransplant == 1)

```

## Does survival differ between the plots with transplants (disturbance) and plots without disturbance? 
```{r}

exist2023_2024 <- surv2023_2024 %>%
  filter(isTransplant == 0) %>% 
  left_join(trans2024, by = c("transectID" = "transectId")) %>%
  filter(!(transectID %in% c(15, 16, 21))) %>%
  mutate(Date = as.Date(date)) %>%
  arrange(Date) %>%
  ## Multiple measurements per year
  mutate(year = format(Date, "%Y")) %>%
  group_by(year, transectID, type_CT_TT, plantID) %>%
  summarise(heightInCentimeters = max(heightInCentimeters, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(surv = case_when(heightInCentimeters > 0 ~ 1,
                          heightInCentimeters == 0 ~ 0)) %>%
  ## All have to be alive in the first year
  filter(!(year == 2023 & surv == 0)) 

ggsave(filename = "C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/Figures/SurvXTransectTreatment.jpg", 

exist2023_2024 %>%
  pivot_wider(id_cols = c(transectID, type_CT_TT, plantID), names_from = year, names_prefix = "Year",
              values_from = surv) %>%
  group_by(transectID, type_CT_TT) %>%
  # filter(is.na(Year2023))
  ## No zeros in 2023, new in 2024  
  summarise(survRate = sum(Year2024[Year2023 == 1], na.rm = TRUE)/sum(Year2023, na.rm = TRUE),
            recruit = length(is.na(Year2023))) %>%
  ## I think we didn't get the data on the existing plants in 2023 for treatment plots 2, 4, 6, or 8
  filter(!is.infinite(survRate)) %>%
ggplot(   aes(type_CT_TT, survRate)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(height = 0, pch = 16)+
  theme_bw() +
  scale_x_discrete(labels = c("CT" = "Control", "TT" = "Transplant")) +
  xlab("Transect") +
  ylab("Percent survival") 
# +
#   geom_hline(yintercept = 0.7889) +
#   geom_hline(yintercept = 0.8877)

, width=120, height=100,units='mm', dpi=300)  



existGLM <- exist2023_2024 %>%
  pivot_wider(id_cols = c(transectID, type_CT_TT, plantID), names_from = year, names_prefix = "Year",
              values_from = surv) %>%
  mutate(surv = case_when( (Year2024 == 1 & Year2023 == 1) ~ 1,
                           # (Year2024 == 1 & is.na(Year2023)) ~ NA,
                           (Year2024 == 0 & Year2023 == 1) ~ 0 )) %>% 
  mutate(recruit = ifelse(is.na(Year2023), ifelse(Year2024 == 1, 1, 0), 0) )

lm1 <- glm(surv ~ type_CT_TT, family = binomial(link = "logit"),data = existGLM)

summary(lm1)

## use the oddsratio package
# odds ratio 
TTor <- or_glm(data = existGLM, 
       model = lm1)

# Log odds
exp(sum(lm1$coefficients))
## Odds ratio
exp(lm1$coefficients[1])/exp(lm1$coefficients[1]) * exp(lm1$coefficients[2])

exp(lm1$coefficients[1])/exp(sum(lm1$coefficients))
## exp(TT) tells us expected increase in the odds of survival for being in a treatment
exp(TTor$oddsratio)/(1+exp(TTor$oddsratio))


## Probability of surviving when in treatment
exp(sum(lm1$coefficients))/(1+exp(sum(lm1$coefficients)))
## Probability of surviving when in control
exp(lm1$coefficients[1])/(1+(exp(lm1$coefficients[1])))

## What happened to data for transects 21 and 32??
exist2023_2024 %>%
  pivot_wider(id_cols = c(transectID, type_CT_TT, plantID), names_from = year, names_prefix = "Year",
              values_from = surv) %>%
  group_by(transectID, type_CT_TT) %>%
  # filter(is.na(Year2023))
  ## No zeros in 2023, new in 2024  
  summarise(survRate = sum(Year2024[Year2023 == 1], na.rm = TRUE)/sum(Year2023, na.rm = TRUE),
            recruit = length(is.na(Year2023))) %>%
  ## I think we didn't get the data on the existing plants in 2023 for treatment plots 2, 4, 6, or 8
  filter(!is.infinite(survRate)) %>%
  filter(survRate == 0)


## Transect 21 is lost!! 
surv2023_2024 %>%
  filter(transectID == 21)

## Looks like only a couple existing, all dead
surv2023_2024 %>%
  filter(transectID == 32)
```

# Add 2025 
```{r}
```




## Survival among transplants
```{r}
surv2023_2024 %>%
  filter(transplantSeason %in% c("Spring", "Fall")) %>% 
  filter(transectID != 16) %>%
  mutate(Date = as.Date(date)) %>%
  arrange(Date) %>%
  mutate(surv = case_when(heightInCentimeters > 0 ~ 1,
                          heightInCentimeters == 0 ~ 0)) 


transplants2025 <- plants2025_updated %>%
  filter(duplicateNew == 0) %>%
  filter(transplantSeason %in% c("Spring", "Fall")) %>% 
  mutate(date = as.Date(date, "%m/%d/%Y")) %>%
  arrange(date) %>%
  mutate(surv = case_when(heightInCentimeters > 0 ~ 1,
                          heightInCentimeters == 0 ~ 0)) %>%
  mutate(id.x = id,
         id.y = id) %>%
  select(c(id.x, plantID, date, heightInCentimeters, basalWidthInCentimeters, reproductiveStatus, numberReproductiveStructures,
           comments, recordedBy, id.y, transectID, x, y, transplantSeason, transectPosition, isTransplant))

## After a few more years of data, might be useful. Not sure why the line for first set of Spring isn't at 1
surv23_25 <- surv2023_2024 %>%
  mutate(date = as.Date(date)) %>%
  bind_rows(transplants2025) %>%
  filter(transplantSeason %in% c("Spring", "Fall")) %>% 
  filter(transectID != 16) %>%
  mutate(year = format(as.Date(date, format = "%Y-%m-%d"), "%Y")) %>%
  arrange(year) %>%
  mutate(surv = case_when(heightInCentimeters > 0 ~ 1,
                          heightInCentimeters == 0 ~ 0)) 


surv23_25 %>%
  ggplot(  aes(year, surv, color = transplantSeason)) +
    geom_point(pch = "|")+
    geom_smooth(method = "glm", 
                method.args = list(family = "binomial"), 
                se = TRUE) +
    scale_color_manual(values = c("darkgreen","burlywood1")) +
  theme_bw()



## Percent survival at each date, zero survival and then greater survival??
plot1 <- surv23_25 %>%
  filter(transplantSeason %in% c("Spring", "Fall")) %>% 
  filter(transectID != 16) %>%
  # mutate(Date = as.Date(date)) %>%
  arrange(date) %>%
  filter(date %in% c("2024-06-24", "2025-06-25")) %>%
  mutate(surv = case_when(heightInCentimeters > 0 ~ 1,
                          heightInCentimeters == 0 ~ 0)) %>% 
  group_by(transplantSeason, date, transectID, recordedBy) %>%
  summarise(PercSurv = sum(surv)/5) %>%
  filter(PercSurv != 0) %>%
  ggplot(  aes(as.factor(date), PercSurv, color = transplantSeason)) +
    geom_boxplot(outlier.shape = NA )+
    geom_point(position = position_jitterdodge(jitter.height = 0.01), pch = 16)+
    scale_color_manual(values = c("green4","hotpink4")) +
    theme_bw()+
    xlab("Survey date")+
    ylab("Percent survival")

ggsave(filename = "C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/Figures/SurvXTreatmentYear.jpg", 
       
       plot1
       
, width=140, height=110,units='mm', dpi=300)  


## How many survived by season
surv2023_2024 %>%
  filter(date == "2024-06-24") %>%
  filter(transectID != 16) 


surv2023_2024 %>%
  filter(transplantSeason %in% c("Spring", "Fall")) %>%  
  mutate(date = case_when(date == "2024-06-25" ~ "2024-06-24",
                          TRUE ~ date)) %>%
  mutate(Date = as.Date(date)) %>%
  arrange(Date) %>%
  mutate(surv = case_when(heightInCentimeters > 0 ~ 1,
                          heightInCentimeters == 0 ~ 0)) %>%
  mutate(date = case_when(date == "2024-06-25" ~ "2024-06-24",
                          TRUE ~ date)) %>%
  filter(date == "2024-06-24") %>%
  ggplot(  aes(transectID, surv, color = transplantSeason)) +
    geom_point(pch = "|")+
    geom_smooth(method = "glm", 
                method.args = list(family = "binomial"), 
                se = TRUE) +
    scale_color_manual(values = c("darkgreen","burlywood1")) +
    theme_bw()

## To 2025
############################### Could be they weren't detectable but still there, could see them again in 2025
## Or we confused a new plant with old one in 2025. 
ggsave(filename = "C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/Figures/SurvXTreatmentYearTransect.jpg", 


surv23_25 %>%
  filter(transplantSeason %in% c("Spring", "Fall")) %>%  
  mutate(surv = case_when(heightInCentimeters > 0 ~ 1,
                          heightInCentimeters == 0 ~ 0,
                          is.na(heightInCentimeters) ~ 0)) %>%
  ggplot(  aes(transectID, surv, color = transplantSeason)) +
    geom_point(pch = "|")+
    geom_smooth(method = "glm", 
                method.args = list(family = "binomial"), 
                se = TRUE) +
    scale_color_manual(values = c("darkgreen","burlywood1")) +
    theme_bw()+
  facet_wrap(~year)


, width=150, height=100,units='mm', dpi=300)  


surv2023_2024 %>%
  filter(transplantSeason %in% c("Spring", "Fall")) %>% 
  mutate(date = case_when(date == "2024-06-25" ~ "2024-06-24",
                          TRUE ~ date)) %>% 
  mutate(Date = as.Date(date)) %>%
  arrange(Date) %>%
  mutate(surv = case_when(heightInCentimeters > 0 ~ 1,
                          heightInCentimeters == 0 ~ 0)) %>%
  filter(!(transplantSeason == "Fall" & Date == "2023-05-31")) %>%
  mutate(date = case_when(date == "2024-06-25" ~ "2024-06-24",
                          TRUE ~ date)) %>%
  filter(date == "2024-06-24") %>%
  ggplot(  aes(transectID, surv)) +
    geom_point(pch = "|")+
    geom_smooth(method = "glm", 
                method.args = list(family = "binomial"), 
                se = TRUE) +
    theme_bw()

## Survival depend on size or reproduction?

survXWhat <- surv2023_2024 %>%
  filter(transplantSeason %in% c("Spring", "Fall")) %>%  
  mutate(surv = case_when(heightInCentimeters > 0 ~ 1,
                          heightInCentimeters == 0 ~ 0)) %>%
  mutate(Survey = case_when(date == "2023-05-31" ~ "aSp23",
                            date == "2023-10-16" ~ "bFa23",
                            date == "2024-06-24" ~ "cSm24")) %>%
  mutate(Obsr = case_when(recordedBy == "Michelle DePrenger-Levin" ~ "MEDL",
                          # recordedBy == "Michelle DePrenger-Levin, Adriana Jacobi" ~ "MEDL_AJ",
                          recordedBy == "Mike Bone, Stanton Schell" ~ "MB_SS",
                          recordedBy == "Arich Fruehling, Leah Veldhuisen" ~ "AF_LV",
                          recordedBy == "Geena Poulter, Michelle DePrenger-Levin" ~ "GP_MEDL",
                          recordedBy == "Michael Guidi, Brooke Washburn, Syed Jalalzai" ~ "MG_BW_SJ")) %>%
  pivot_wider(names_from = c(Survey, Obsr),
              values_from = c(heightInCentimeters, basalWidthInCentimeters, reproductiveStatus),
              id_cols = c(plantID, transectID, transplantSeason)) %>%
  relocate(heightInCentimeters_bFa23_MEDL, .after = heightInCentimeters_aSp23_MEDL) %>%
  unite(InitialHeight, heightInCentimeters_aSp23_MEDL:heightInCentimeters_bFa23_MEDL, sep = ",") %>%
  # Split and keep the first
  mutate(InitialHeight = case_when(transplantSeason == "Spring" ~ sub(",.*", "", InitialHeight),
                                   transplantSeason == "Fall" ~ sub(".*,", "", InitialHeight))) %>%
  mutate(InitialHeight = as.numeric(InitialHeight)) %>%
  ## Width!
  relocate(basalWidthInCentimeters_bFa23_MEDL, .after = basalWidthInCentimeters_aSp23_MEDL) %>%
  unite(InitialbasalWidth, basalWidthInCentimeters_aSp23_MEDL:basalWidthInCentimeters_bFa23_MEDL, sep = ",") %>%
  # Split and keep the first
  mutate(InitialbasalWidth = case_when(transplantSeason == "Spring" ~ sub(",.*", "", InitialbasalWidth),
                                   transplantSeason == "Fall" ~ sub(".*,", "", InitialbasalWidth))) %>%
  mutate(InitialbasalWidth = as.numeric(InitialbasalWidth)) %>%
  ## Initial reproductive status
  unite(InitialRepro, c(reproductiveStatus_aSp23_MEDL,reproductiveStatus_bFa23_MEDL), sep = ",") %>%
  mutate(InitialRepro = case_when(transplantSeason == "Spring" ~ sub(",.*", "", InitialRepro),
                                   transplantSeason == "Fall" ~ sub(".*,", "", InitialRepro))) %>%
  relocate(InitialRepro, .after = InitialHeight) %>%
  ## Get survival across different people recording
  rowwise() %>%
  # mutate(maxHeight = max(across(heightInCentimeters_cSm24_AF_LV:heightInCentimeters_cSm24_MG_BW_SJ), na.rm = TRUE)) %>%
  # select(c(plantID:InitialHeight, maxHeight))
  # 
  mutate(surv = if_else(max(across(heightInCentimeters_cSm24_AF_LV:heightInCentimeters_cSm24_MG_BW_SJ), na.rm = TRUE) > 0,
                        1, 0)) %>%
  relocate(surv, .after = InitialHeight) %>%
  relocate(InitialbasalWidth, .after = InitialHeight)
  


ggplot(survXWhat,  aes(InitialHeight, surv, color = transplantSeason)) +
  geom_point(pch = "|")+
  geom_smooth(method = "glm", 
              method.args = list(family = "binomial"), 
              se = TRUE) +
  theme_bw()

ggplot(survXWhat,  aes(InitialbasalWidth, surv, color = transplantSeason)) +
  geom_point(pch = "|")+
  geom_smooth(method = "glm", 
              method.args = list(family = "binomial"), 
              se = TRUE) +
  theme_bw()

```

```{r}

#### Overall Survival
survXWhat %>%
  group_by(transectID, transplantSeason) %>%
  summarise(percsurv = sum(surv)/n()) %>%
ggplot(   aes(transplantSeason, percsurv, fill = transplantSeason)) +
  geom_boxplot() +
  geom_jitter(position = position_jitterdodge(jitter.height = 0.01), pch = 16) +
  theme_bw() +
  scale_x_discrete(labels = c("0" = "Vegetative", "1" = "Reproductive")) +
  xlab("Initial Reproductive Status") +
  ylab("Percent survival") +
  scale_fill_manual("",values = c("Spring"="darkgreen","Fall"="burlywood1"))

survXWhat %>%
  group_by(transectID, transplantSeason) %>%
  summarise(percsurv = sum(surv)/n()) %>%
  group_by(transplantSeason) %>%
  summarise(MeanSurv = mean(percsurv),
            SDsurv = sd(percsurv)) 

survXWhat %>%
  group_by(transectID) %>%
  summarise(percsurv = sum(surv)/n()) %>%
  ungroup() %>%
  summarise(avgSurv = mean(percsurv))


#### Survival spatially
survXWhat %>%
  group_by(transectID, transplantSeason) %>%
  summarise(percsurv = sum(surv)/n()) %>%
ggplot(   aes(transectID, percsurv, color = transplantSeason)) +
    geom_point(pch = "|")+
    stat_smooth(method="glm",  method.args = list(family = "binomial"), se=TRUE, alpha = 0.15)+
  theme_bw() +
  xlab("Transects/n(South to North)") +
  ylab("Percent survival") +
  scale_color_manual("",values = c("Spring"="darkgreen","Fall"="burlywood1"))




############## Figure 2: survival by initial reproductive status
ggsave(filename = "C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/Figures/SurvXSeason&ReproStatus.jpg", 

survXWhat %>%
  group_by(transectID, transplantSeason, InitialRepro) %>%
  summarise(percsurv = sum(surv)/n()) %>%
ggplot(   aes(InitialRepro, percsurv, fill = transplantSeason)) +
  geom_boxplot() +
  geom_jitter(position = position_jitterdodge(jitter.height = 0.01), pch = 16) +
  theme_bw() +
  scale_x_discrete(labels = c("0" = "Vegetative", "1" = "Reproductive")) +
  xlab("Initial Reproductive Status") +
  ylab("Percent survival") +
  scale_fill_manual("",values = c("darkgreen","burlywood1"))

, width=120, height=100,units='mm', dpi=300)  



survXWhat %>%
  group_by(transectID, transplantSeason, InitialRepro) %>%
  summarise(percsurv = sum(surv)/n()) %>%
  group_by(transplantSeason, InitialRepro) %>%
  summarise(MeanSurv = mean(percsurv),
            SDsurv = sd(percsurv)) 
```

AIC compare all the reasons for survival of transplants   
```{r}
## What influences survival from 2023 to 2024

lm1 <- glm(surv ~ InitialHeight*transplantSeason, data = survXWhat, family = binomial(link = "logit"))
lm2 <- glm(surv ~ as.factor(transectID)*transplantSeason, data = survXWhat, family = binomial(link = "logit")) # Spatially?
lm3 <- glm(surv ~ InitialbasalWidth*transplantSeason, data = survXWhat, family = binomial(link = "logit"))
lm4 <- glm(surv ~ InitialRepro*transplantSeason, data = survXWhat, family = binomial(link = "logit"))
lm5 <- glm(surv ~ transplantSeason, data = survXWhat, family = binomial(link = "logit"))
lm6 <- glm(surv ~ as.factor(transectID), data = survXWhat, family = binomial(link = "logit")) # Spatially not by season

  
lm.list <- list(lm1,lm2,lm3,lm4, lm5,lm6)
lm.names <- as.character(unlist(lapply(lm.list,formula)))
(lm.results <- aictab(lm.list, modnames=lm.names))
#evidence ratio 
for(i in 2:length(lm.list)){
  print(exp(0.5*lm.results$Delta_AICc[i]))
}

evidence(lm.results)


```





## Survival by initial density of plants
```{r}

### Calculate initial density in fall 2023 and then bind to survXWhat

InitDensity <- surv2023_2024 %>%
  filter(isTransplant == 0) %>%
  group_by(transectID, date, recordedBy) %>%
  summarise(abundance = n()) %>%
  ## Average over duplicate observations
  group_by(transectID, date) %>%
  summarise(avgAbund = mean(abundance)) %>%
  filter(date == "2024-06-24")

############## survival by density aka abundance
# ggsave(filename = "C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/Figures/SurvXSeason&ReproStatus.jpg", 

survXWhat %>%
  left_join(InitDensity, by = "transectID") %>%
  group_by(transectID, transplantSeason, avgAbund) %>%
  summarise(percsurv = sum(surv)/n()) %>%
ggplot(   aes(avgAbund, percsurv, color = transplantSeason)) +
    geom_point(pch = "|")+
    stat_smooth(method="glm",  method.args = list(family = "binomial"), se=TRUE)+
  theme_bw() +
  xlab("Conspecific abundance") +
  ylab("Percent survival") +
  scale_color_manual("",values = c("darkgreen","burlywood1"))

# , width=120, height=100,units='mm', dpi=300)  



survXWhat %>%
  group_by(transectID, transplantSeason, InitialRepro) %>%
  summarise(percsurv = sum(surv)/n()) %>%
  group_by(transplantSeason, InitialRepro) %>%
  summarise(MeanSurv = mean(percsurv),
            SDsurv = sd(percsurv))



```



```{r}
## Need initial height
surv_wide_perfectdetection <- surv2023_2024 %>%
  mutate(date = case_when(date == "2024-06-25" ~ "2024-06-24",
                          TRUE ~ date)) %>%
  mutate(Date = as.Date(date)) %>%
  arrange(Date) %>%
  mutate(date2 = gsub("[[:punct:]]", "", surv2023_2024$date))  %>%
  pivot_wider(names_from = c(date2),
              values_from = heightInCentimeters,
              id_cols = c(plantID,transplantSeason,transectID, Block),
              names_prefix = "Date",
               ,values_fn = function(x) max(x,na.rm = TRUE)  ## Because why? why are there double for the same date and observers?
              ) %>%
  arrange(transectID,Block,transplantSeason) %>%
  mutate(InitialHeight = case_when(transplantSeason == "Spring" ~ Date20230531,
                                   transplantSeason == "Fall" ~ Date20231016))

## What?!?! Now there are more measurements for fall in the spring! that's wrong!!! 
surv_wide_perfectdetection %>%
  filter(transplantSeason %in% c("Spring","Fall")) %>%
  filter(transplantSeason == "Fall" & !is.na(Date20230531))

surv_wide_perfectdetection %>%
  filter(transplantSeason %in% c("Fall","Spring")) %>%
  mutate(surv = case_when('2024-06-24' > 0 ~ 1,
                          '2024-06-24' == 0 ~ 0)) %>%
  group_by(transectID, Block, transplantSeason) %>%
  summarise(percSurv = sum(surv)/5)
  ggplot(  aes(transplantSeason, surv))

```



## Survival by transplant time  
```{r}

surv_transplants <- surv2023_2024 %>%
  filter(transplantSeason %in% c("Spring", "Fall")) %>%  
  mutate(InitialHeight = case_when(transplantSeason=="Spring" ~ heightInCentimeters[date == "2023-05-31"],
                                   transplantSeason=="Fall" ~ heightInCentimeters[date == "2023-10-16"])) %>%
  mutate(Date = as.Date(date)) %>%
  arrange(Date) %>%
  mutate(surv = case_when(heightInCentimeters > 0 ~ 1,
                          heightInCentimeters == 0 ~ 0)) %>%
  filter()


glm.transplantSeason <- glm()



glm.list <- list(brnull,br1a,br2,br3,br1b,br1,br2a, br3a)
glm.names <- as.character(unlist(lapply(glm.list,formula)))
(glm.results <- aictab(glm.list, modnames=glm.names))

```


### Run with rmark  
```{r}

```





### Examples for true models

Gimenez 2020 <https://www.youtube.com/watch?v=VR8qdNvCaGk>     
phi(.), p(.)   
```{r}

# Likelihood
for(i in 1:nind){
  
  # Define latent state at first capture
  z[i,f[i]] <- 1  # vector of when first encounter happened, always alive when first detected
  
  for(t in (f[i]+1):n.occasions){  # loop over time from the second encounter onward
    # State process
    z[i,t] ~ dbern(phi * z[i,t-1]) # will be alive on first encounter, then Bernoulli
    
    # Obervation process
    y[i,t] ~ dbern(p * z[i,t]) # true state times prob of detection
    
    } # t time
  
} # loop over all i individuals

phi ~ dunif(0,1) # prior for survival
p ~ dunif(0,1) # Prior for recapture




```

CJS phi(t), p(t)   
```{r}

# Likelihood
for(i in 1:nind){
  
  # Define latent state at first capture
  z[i,f[i]] <- 1  # vector of when first encounter happened, always alive when first detected
  
  for(t in (f[i]+1):n.occasions){  # loop over time from the second encounter onward
    # State process
    z[i,t] ~ dbern(phi[t-1] * z[i,t-1]) # will be alive on first encounter, then Bernoulli, survival for each time interval
    
    # Obervation process
    y[i,t] ~ dbern(p[t-1] * z[i,t]) # true state times prob of detection, detection for each time interval
    
    } # t time
  
} # loop over all i individuals

## Need prior for each occasion
for(t in 1:n.occasions-1){
  phi[t] ~ dunif(0,1) # prior for survival
  p[t] ~ dunif(0,1) # Prior for recapture
}




```


Differences among groups, individual covariates     
```{r}


# Likelihood
for(i in 1:nind){
  
  # Define latent state at first capture
  z[i,f[i]] <- 1  # vector of when first encounter happened, always alive when first detected
  
  for(t in (f[i]+1):n.occasions){  # loop over time from the second encounter onward
    # State process
    z[i,t] ~ dbern(phi[i,t-1] * z[i,t-1]) # will be alive on first encounter, then Bernoulli, survival for each time interval
    
    # Obervation process
    y[i,t] ~ dbern(p[i,t-1] * z[i,t]) # true state times prob of detection, detection for each time interval
    
    } # t time
  
} # loop over all i individuals

## Need prior for each occasion
for(i in 1:nind){
  for(t in 1:(n.occasions-1)){  # in interval from t to t+1
    phi[i,t] ~ dunif(0,1) # prior for survival
    p[i,t] ~ dunif(0,1) # Prior for recapture
  }
}

```

Apply constraints on parameters, phi(.), p(.)   
```{r}

# Priors and constraints 
for(i in 1:nind){
  for(t in 1:(n.occasions-1)){
    phi[i,t] <- mean.phi   # not a 'mean' but a constant
    p[i,t] <- mean.p       # not a 'mean' but a constant
  }
}  

# Prior  
mean.phi ~ dunif(0,1)
mean.p ~ dunif(0,1)


```



Simulate capture-history matrix   
```{r}
n.occasions <- 6
marked <- rep(50, n.occasions-1)   # Annual number of newly marked individuals
phi <- rep(0.65, n.occasions-1)
p <- rep(0.4, n.occasions-1)

# Define matrices with survival and recapture probs
PHI <- matrix(phi, ncol = n.occasions-1, nrow = sum(marked))
P <- matrix(p, ncol = n.occasions-1, nrow=sum(marked))

simul.cjs <- function(PHI, P, marked){
  n.occasions <- dim(PHI)[2] + 1
  CH <- matrix(0, ncol = n.occasions, nrow = sum(marked))
  
  # Define a vector with the occasion of marking
  mark.occ <- rep(1:length(marked), marked[1:length(marked)])
  # Fill the CH matrix 
  for(i in 1:sum(marked)){
    CH[i, mark.occ[i]] <- 1     # First is a 1
    if(mark.occ[i] == n.occasions) next
      for(t in (mark.occ[i]+1):n.occasions){
        # Bernoulli trial for survival
        sur <- rbinom(1,1,PHI[i,t-1])
        if(sur==0) break   # If dead, move to next individual
        # Bernoulli for recapture
        rp <- rbinom(1,1,P[i,t-1])
        if(rp==1) CH[i,t] <- 1  # If you recaptured, it gets a '1'
      } # t
  } # i
return(CH)
  }

## Simulate!
sim1 <- simul.cjs(PHI, P, marked)

# Create vector with occasion of marking  
get.first <- function(x) min(which(x != 0))
f <- apply(sim1, 1, get.first)  ## Apply across rows


## BUGS  
SimulCJS <- 
  paste("
model {
  
  # Constraints
  for(i in 1:nind){
    for(t in 1:(n.occasions-1)){
      phi[i,t] <- mean.phi
      p[i,t] <- mean.p
    }
  }
  
  # Priors
  mean.phi ~ dunif(0,1)
  mean.p ~ dunif(0,1)
  
  # Likelihood
  for(i in 1:nind){
  
    # Define latent state at first capture
    z[i,f[i]] <- 1
    for(t in (f[i]+1):n.occasions){
    
      # State process
      z[i,t] ~ dbern(phi[i,t-1] * z[i, t-1])
      
      # Observation process
      y[i,t] ~ dbern(p[i,t-1] * z[i,t])
    }
  }
}

")

writeLines(SimulCJS, "SimulCJS.jags")

jags.data <- list(y = sim1,
                  f= f,
                  nind = nrow(sim1),
                  n.occasions = ncol(sim1))

# Set 'good' initial values
z.inits <- function(ch){
  state <- ch
  state[state==0] <- 1
  get.first <- function(x){ min(which(x != 0)) }
  f <- apply(ch, 1, get.first)
    for(i in 1:nrow(ch)){
      state[i, 1:f[i]] <- NA
    }
  return(state)
}

inits <- function(){list(mean.phi = runif(1,0,1),
                         mean.p = runif(1,0,1),
                         z = z.inits(sim1))}   # The latent states (alive, dead) also need initial values
  # Could monitor the latent states (if we have missing data)

parameters <- c("mean.phi","mean.p")

# MCMC settings
ni <- 1000
nt <- 1
nb <- 500
nc <- 3

cjs.c.c <- jags(data = jags.data,
                inits = inits,
                parameters.to.save = parameters,
                model.file = "SimulCJS.jags",
                n.chains = nc,
                n.thin = nt,
                n.iter = ni, n.burnin = nb,
                working.directory = getwd())

print(cjs.c.c, digits = 3)

## This is around 50 minutes

# save(cjs.c.c, file = )
```



Second part <https://www.youtube.com/watch?v=vYWj50Sj-q4>  
Random time effects
```{r}

n.occasions >- 12
marked <- rep(50, n.occasions-1)  ## Annual number of newly marked individuals
mean.phi <- 0.65
sigma2.phi <- 1
p <- rep(0.4, n.occasions-1)

## Annual survival probabilities on logistic scale
logit.phi <- rnorm(n.occasions-1, qlogis(mean.phi), sigma2.phi^0.5)
logit.phi
phi <- plogis(logit.phi)
phi
```


https://www.youtube.com/watch?v=vYWj50Sj-q4 
stopped at Minute 29



Tomst data logger  
Put in issue on GitHub
```{r}

# install.packages("myClim")


tms.f <- mc_read_files(c("C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_lolly_TOMSTdatalogger/data_95135950_2024_06_24_0.csv",
                         "C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_lolly_TOMSTdatalogger/data_95135951_2024_06_24_0.csv",
                         "C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_lolly_TOMSTdatalogger/data_95135952_2024_06_24_0.csv"),
                       dataformat_name = "TOMST", silent = TRUE)


tms.fs <- mc_read_files(paths = "C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_lolly_TOMSTdatalogger",
                        dataformat_name = "TOMST")

```

