---
title: "FuzzyMatch PlantIDs"
author: "Michelle DePrenger-Levin"
date: "2024-06-27"
output: html_document
---

Data entry, matching for new PlantIDs
```{r}

rm(list=ls()) 
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)



```


Read data on 2024-06-24 and 2024-06-25
Three teams one day one; four teams on day two  
```{r}

pepe1 <- read.csv("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/PePeTransplant/20240625_penstemon-penlandii_ArichLeah_csv.csv",
                  skip = 2,
                  header = TRUE)
pepe2 <- read.csv("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/PePeTransplant/20240625_penstemon-penlandii_GeenaMEDL_csv.csv",
                  skip = 2,
                  header = TRUE)
pepe3 <- read.csv("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/PePeTransplant/20240625_penstemon-penlandii_Mike Bone & Stanton Schell_csv.csv",
                  skip = 2,
                  header = TRUE)
pepe4 <- read.csv("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/PePeTransplant/20240625_penstemon-penlandii_Michael Guidi Brooke Washburn Syed Jalalzai_csv.csv",
                  skip = 2,
                  header = TRUE)
pepe5 <- read.csv("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/PePeTransplant/20240624_penstemon-penlandii_MBoneStanton_csv.csv",
                  skip = 2,
                  header = TRUE)
pepe6 <- read.csv("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/PePeTransplant/20240624_penstemon-penlandii_ArichLeah_csv.csv",
                  skip = 2,
                  header = TRUE)
pepe7 <- read.csv("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/PePeTransplant/20240624_penstemon-penlandii_GeenaMEDL_csv.csv",
                  skip = 2,
                  header = TRUE)

```


Merge, only keep sites with data collected    
Transect 15 is new, with new start and end
```{r}

pepe2024 <- pepe1 %>%
  bind_rows(pepe2) %>%
  bind_rows(pepe3) %>%
  bind_rows(pepe4) %>%
  bind_rows(pepe5) %>%
  bind_rows(pepe6) %>%
  bind_rows(pepe7) %>%
  filter(transectId != "transectId") %>%
  filter(Date != "") %>%
  filter(end5mTagNum != "new tag and new line") %>%
  mutate(across(transectId:numRep_CurrentYear, as.numeric)) %>%
         # = as.numeric(transectId),
         # originTagNum = as.numeric(originTagNum),
         # end5mTagNum = as.numeric(end5mTagNum),
         # plantID = as.numeric(plantID),
         # x = as.numeric(x),
         # y = as.numeric(y))
  ## Fill in missing TransectID, starting and ending tag numbers, defaults to replacing missing data from top to bottom
  fill(transectId) %>%
  fill(originTagNum) %>%
  fill(end5mTagNum) %>%
  filter(!(transectId == 1 & Recorded.by == "Geena Poulter, Michelle DePrenger-Levin"))


## Loop across anything without a plantID, look at first row, find within +/-, do max plantID+1
for(i in 1:nrow(pepe2024)){
  if(is.na(pepe2024$plantID[i])){
    for(n in which(is.na(pepe2024$plantID)) )
  } # end if statement with new plant
}


## Sort by Y and X within each
checknew <- pepe2024 %>%
  arrange(transectId, y, x) %>%
  group_by(transectId, Date, Recorded.by) %>%
  summarise(max(plantID, na.rm = TRUE))
  ### Use plantID; when new (no plantID) then add a number to each team's records, see if in order they match
  mutate(newID = case_when(!is.na(plantID) ~ plantID,
                           is.na(plantID) ~ seq(from = max(plantID, na.rm = TRUE), to = length(is.na(plantID))))) %>%  
  pivot_wider(id_cols = c(transectId, originTagNum, end5mTagNum, newID), names_from = Recorded.by, 
              values_from = c(x:y)) 

pepe2024 %>%
  ## Match X and Y within a range for new plants, enter average of X and Y as final coordinates
  mutate(Xnew = x, Ynew = y, PlantIDnew = plantID) 
  group_by(transectID) %>%
  mutate(PlantIDnew = case_when(plantID == "" & ))
  ## Sort by PlantID and autofill the next numbers for new plants
  


```



Need to match X and Y that are close to each other, group by those and then take the average for the final X and Y   
Plot 37 was only done once but half by Stanton Schell and Syed Jalalzi and half by MEDL and Geena P. 
```{r}




```

