---
title: "FuzzyMatch PlantIDs"
author: "Michelle DePrenger-Levin"
date: "2024-06-27"
output: html_document
---

Data entry, matching for new PlantIDs
```{r}

rm(list=ls()) 
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)



```


Read data on 2024-06-24 and 2024-06-25
Three teams one day one; four teams on day two  
```{r}

pepe1 <- read.csv("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/PePeTransplant/20240625_penstemon-penlandii_ArichLeah_csv.csv",
                  skip = 2,
                  header = TRUE)
pepe2 <- read.csv("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/PePeTransplant/20240625_penstemon-penlandii_GeenaMEDL_csv.csv",
                  skip = 2,
                  header = TRUE)
pepe3 <- read.csv("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/PePeTransplant/20240625_penstemon-penlandii_Mike Bone & Stanton Schell_csv.csv",
                  skip = 2,
                  header = TRUE)
pepe4 <- read.csv("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/PePeTransplant/20240625_penstemon-penlandii_Michael Guidi Brooke Washburn Syed Jalalzai_csv.csv",
                  skip = 2,
                  header = TRUE)
pepe5 <- read.csv("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/PePeTransplant/20240624_penstemon-penlandii_MBoneStanton_csv.csv",
                  skip = 2,
                  header = TRUE)
pepe6 <- read.csv("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/PePeTransplant/20240624_penstemon-penlandii_ArichLeah_csv.csv",
                  skip = 2,
                  header = TRUE)
pepe7 <- read.csv("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/PePeTransplant/20240624_penstemon-penlandii_GeenaMEDL_csv.csv",
                  skip = 2,
                  header = TRUE)

```


Merge, only keep sites with data collected    
Transect 15 is new, with new start and end
```{r}

pepe2024 <- pepe1 %>%
  bind_rows(pepe2) %>%
  bind_rows(pepe3) %>%
  bind_rows(pepe4) %>%
  bind_rows(pepe5) %>%
  bind_rows(pepe6) %>%
  bind_rows(pepe7) %>%
  filter(transectId != "transectId") %>%
  filter(Date != "") %>%
  filter(end5mTagNum != "new tag and new line") %>%
  mutate(across(transectId:numRep_CurrentYear, as.numeric)) %>%
         # = as.numeric(transectId),
         # originTagNum = as.numeric(originTagNum),
         # end5mTagNum = as.numeric(end5mTagNum),
         # plantID = as.numeric(plantID),
         # x = as.numeric(x),
         # y = as.numeric(y))
  ## Fill in missing TransectID, starting and ending tag numbers, defaults to replacing missing data from top to bottom
  fill(transectId) %>%
  fill(originTagNum) %>%
  fill(end5mTagNum) %>%
  filter(!(transectId == 1 & Recorded.by == "Geena Poulter, Michelle DePrenger-Levin")) 

## A few errors
pepe2024 %>%
  filter(is.na(x))

## Add a dummy X for the one record
rownum <- which(is.na(pepe2024$x) & !is.na(pepe2024$height_CurrentYear))
pepe2024$notes_CurrentYear[is.na(pepe2024$x) & !is.na(pepe2024$height_CurrentYear)] <- "Missing X, added placeholder 0.5"
pepe2024$x[is.na(pepe2024$x) & !is.na(pepe2024$height_CurrentYear)] <- 0.5
pepe2024[rownum,]

pepe2024 <- pepe2024 %>%
  filter(!is.na(x))
## Check Y
pepe2024 %>%
  filter(is.na(y))

## Loop across anything without a plantID, look at first row, find within +/-, do max plantID+1
for(i in 1:nrow(pepe2024)){
  
  seenbyboth <- FALSE
  
  if(is.na(pepe2024$plantID[i])){
    NAs <- which(is.na(pepe2024$plantID))
    
    ### Until you get to the last one
    if(length(NAs) > 1) {
      for(n in 2:length(NAs)){
        if(pepe2024$transectId[i] == pepe2024$transectId[NAs[n]] &
           ## absolute value of the difference between them x < 0.05; y < 0.03
           abs(pepe2024$x[i] - pepe2024$x[NAs[n]]) < 0.05 &
           abs(pepe2024$y[i] - pepe2024$y[NAs[n]]) < 0.03 & 
           pepe2024$Recorded.by[i] != pepe2024$Recorded.by[NAs[n]]){
          
          pepe2024$plantID[i] <- pepe2024$plantID[NAs[n]] <- (max(pepe2024$plantID, na.rm = TRUE) + 1)
          seenbyboth <- TRUE
          
          } # end matching if statement
      } # for loop checking for matches among other records
      
      ## If only seen by one, then assign unique plantid
      if(!seenbyboth)  pepe2024$plantID[i] <- (max(pepe2024$plantID, na.rm = TRUE) + 1)
      
      # Up to the last one, when nothing to compare to, then give next plantID number
      } else {
      pepe2024$plantID[i] <- (max(pepe2024$plantID, na.rm = TRUE) + 1) 
      }
    
  } # end if statement with new plant
} # end loop through all records

## Sort by Y and X within each
checknew <- pepe2024 %>%
  arrange(transectId, y, x) %>%
  group_by(transectId, Date, Recorded.by) %>%
  pivot_wider(id_cols = c(transectId, originTagNum, end5mTagNum, plantID), names_from = Recorded.by, 
              values_from = c(x:y),
              values_fn = function(x) paste(x, collapse = "; ")) 

checknew %>%
  write.csv("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/PePeTransplant/checkMatchingXY.csv")

pepe2024 %>%
  ## Match X and Y within a range for new plants, enter average of X and Y as final coordinates
  mutate(Xnew = x, Ynew = y, PlantIDnew = plantID) 
  group_by(transectID) %>%
  mutate(PlantIDnew = case_when(plantID == "" & ))
  ## Sort by PlantID and autofill the next numbers for new plants
  


```



Need to match X and Y that are close to each other, group by those and then take the average for the final X and Y   
Plot 37 was only done once but half by Stanton Schell and Syed Jalalzi and half by MEDL and Geena P. 
```{r}




```

