---
title: "Pepe_transplant"
author: "Michelle DePrenger-Levin"
date: "2023-08-29"
output: html_document
---

---
```{r}

rm(list=ls()) 
library(R2jags)
library(runjags)
library(mcmcplots)
library(boot)
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)

library(RMark)
```



```{r}
pepe.spring <- read.csv("C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/2023_Penstemon-penlandii_Annualreport/2023Spring_Penstemon-penlandii.csv")

pepe.fall <- read.csv("C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/2023_Penstemon-penlandii_Annualreport/20231016Fall_Penstemon-penlandii_dataentry_2.csv")

## need to confirm 
pepe.20230626 <- read.csv("C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/2023_Penstemon-penlandii_Annualreport/20230626Spring_Penstemon-penlandii.csv")

## Need control plot data
pepe.20231016_control <- read.csv("C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/2023_Penstemon-penlandii_Annualreport/20231016_ControlTransects_dataentry.csv")

## Data from Fall 2023-10-16 and 2320-10-17 
pepe.springfall <- read.csv("C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/2023_Penstemon-penlandii_Annualreport/20231016Spring_Penstemon-penlandii_fall.csv")

```


## All the data
```{r}
pepe.all.2023 <- read.csv("C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/Data/Penstemon-penlandii_data.csv")

pepe.transects <- read.csv("C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/Data/Penstemon-penlandii_Transects.csv")


## Add the pepe.springfall fall repeat data of all the spring transplants
falltomerge <- pepe.springfall %>%
  left_join(pepe.transects) %>%
  select(c(TransectID, Block, Season, Transect_position, PlantID,Date,  Height.cm.1, BasalWidth.cm.1, Repro.1, NumRep.1, Comments.1, X, Y)) 

names(falltomerge) <- names(pepe.all.2023)
falltomerge$Date <- 20231016

# XY wrong for Trans10, position 3 and Trans36, position 10
pepe.all.2023$X[pepe.all.2023$Transect_position == 3 & 
                  pepe.all.2023$TransectID == 10 & 
                  !is.na(pepe.all.2023$Transect_position)] <- 0.25

pepe.all.2023$X[pepe.all.2023$Transect_position == 10 & 
                  pepe.all.2023$TransectID == 36 & 
                  !is.na(pepe.all.2023$Transect_position)] <- 0.44
pepe.all.2023$Y[pepe.all.2023$Transect_position == 10 & 
                  pepe.all.2023$TransectID == 36 & 
                  !is.na(pepe.all.2023$Transect_position)] <- 4.9
pepe.all.2023$Y[pepe.all.2023$Transect_position == 9 & 
                  pepe.all.2023$TransectID == 12 & 
                  !is.na(pepe.all.2023$Transect_position)] <- 4.25

pepe.all.2023$Y[pepe.all.2023$Transect_position == 4 & 
                  pepe.all.2023$TransectID == 14 & 
                  !is.na(pepe.all.2023$Transect_position)] <- 2.55
pepe.all.2023$Y[pepe.all.2023$Transect_position == 9 & 
                  pepe.all.2023$TransectID == 18 & 
                  !is.na(pepe.all.2023$Transect_position)] <- 4.75

pepe.all.2023$Y[pepe.all.2023$Transect_position == 10 & 
                  pepe.all.2023$TransectID == 32 & 
                  !is.na(pepe.all.2023$Transect_position)] <- 4.45


pepe.all.2023$X[pepe.all.2023$Transect_position == 6 & 
                  pepe.all.2023$TransectID == 36 & 
                  !is.na(pepe.all.2023$Transect_position)] <- 0.60
pepe.all.2023$Y[pepe.all.2023$Transect_position == 6 & 
                  pepe.all.2023$TransectID == 36 & 
                  !is.na(pepe.all.2023$Transect_position)] <- 2.75



## existing and transplanted
# Add a unique ID for each plant
# Has only one observation per plant
pepe.tags <- pepe.all.2023 %>%
  left_join(pepe.transects, by = c("TransectID","Block")) %>%
  select(TransectID:Block, Season:Transect_position, X,Y, Type_CT_TT) %>%
  group_by(TransectID,Block,X,Y) %>% #,Season,Transect_position) %>%
  summarise(PlantID = cur_group_id()) %>%
  # mutate(PlantID = cur_group_id()) %>%
  arrange(TransectID, Block, Y, X)

# Correctly gave each plant an ID number
pepe.tags %>%
  group_by(PlantID) %>%
  summarise(n = n()) %>%
  filter(n > 1)

## Save 'tags' for MySQL  
# write.csv(pepe.tags, file = "C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/Data/Penstemon-penlandii_tags_MySQL.csv")


tags_MySQL <- read.csv("C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/Data/Penstemon-penlandii_tags_MySQL.csv")

tags_MySQL %>%
  left_join(pepe.all.2023, by = c("TransectID", "Block", "X", "Y")) %>%
  select(X.1:Transect_position) %>%
  rename(PlantID = PlantID.x) %>%
  rename(id = X.1) %>%
  rename(x = X) %>%
  rename(y = Y) %>%
  write.csv(  file = "C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/Data/Penstemon-penlandii_tags_MySQL_season&position.csv")

## Save 'tags' for MySQL with Season and Transect_position  
# write.csv(pepe.tags, file = "C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/Data/Penstemon-penlandii_tags_MySQL_season&position.csv")
  

## Backup datasheets to print
pepe.datasheets <- pepe.all.2023 %>%
  left_join(pepe.transects, by = c("TransectID","Block")) %>%
  select(TransectID:Block, Height.cm, Transect_position, X,Y) %>%
  arrange(TransectID, Block, Y, X) 

write.csv(pepe.datasheets, file = "C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/Data/Penstemon-penlandii_datasheets.csv")



pepe.all.2023v2 <- pepe.all.2023 %>%
  # bind_rows(falltomerge) %>%
  select(!"PlantID") %>% 
  left_join(pepe.tags) 
# %>% #, by = c("TransectID","Block","PlantID", "X","Y","Season","Transect_position")) %>%
#   distinct(TransectID, Block, X,Y, #PlantID,
#            .keep_all = TRUE)
  
pepe.all.2023v2 %>%
  filter(is.na(PlantID))


```


# Annual data   
```{r}


### Save a list of data for each occurrence (1st or second and date) for double observer 
# need PlantID, Height, BasalWidth, FlYN, NumFlFr, comments, observer, date  
## Observer should be the one finding and measuring the plants. Setting to MEDL for 2023 as place holder, can enter real observers for all later
pepe.all.2023v2 %>%
  select(c(PlantID, Date:Comments)) %>%
  mutate(Observer = "Michelle DePrenger-Levin") %>%
  write.csv("C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/Data/Penstemon-penlandii_data_MySQL.csv")
  


spring2fall2023 <- pepe.all.2023 %>%
  filter(Season == "Spring") %>%
  pivot_wider(id_cols = c(TransectID, Block, Season, Transect_position, X,Y, PlantID), 
                          names_from = Date, values_from = Height.cm,  names_prefix = "Height") %>%
  select(TransectID:PlantID, Height20230531, Height20231016)


pepe.all.2023v2 %>%
  # select(c(TransectID:Height.cm,X:PlantID)) %>%
  mutate(Date = as.numeric(Date)) %>%
  arrange(Block, TransectID, Y,X, Transect_position, Date) %>%
  pivot_wider(id_cols = c(TransectID, Block, Season, Transect_position, X,Y, PlantID), 
                          names_from = Date, values_from = Height.cm,  names_prefix = "Height") %>% # , values_fill = 0
  # pivot_wider(names_from = Date, values_from = c(Height.cm, BasalWidth.cm, Repro, NumRep, Comments), values_fill = 0) %>%
  select(TransectID:PlantID, Height20230531, Height20231016) %>%
  write.csv("C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/DataSheets_backup/2024PePeDatasheets.csv")
  

pepe.all.2023v2 %>%
  select(c(TransectID:Height.cm,X:PlantID)) %>%
  mutate(Date = as.numeric(Date)) %>%
  arrange(Block, TransectID, Y,X, Transect_position, Date) %>%
  pivot_wider(names_from = Date, values_from = Height.cm, names_prefix = "Date") %>% # values_fill = 0) %>%
  filter(!is.na(Transect_position)) %>%
  filter(Season == "Spring" & !is.na(Date20231016))

```


Format for datasheets (2023)
```{r}


transBlock <- data.frame(TransectID = 1:40, Block = rep(1:20, each = 2))

pepe.spring <- pepe.spring %>%
  mutate(Date = as.Date("2023-05-31")) %>%
  inner_join(transBlock) %>%
  arrange(Block, TransectID)

Transect_positions <- data.frame(Transect_position = 1:10, xy = c("0.25, 0.25",
                                                                  "0.75, 0.75",
                                                                  "0.25, 1.25",
                                                                  "0.75, 1.75",
                                                                  "0.25, 2.25",
                                                                  "0.75, 2.75",
                                                                  "0.25, 3.25",
                                                                  "0.75, 3.75",
                                                                  "0.25, 4.25",
                                                                  "0.75, 4.75"))
  


## When only spring
pepe.annual <- pepe.spring %>%
  mutate(RandPlant = PlantID) %>%
  complete(nesting(TransectID, Block), Transect_positions) %>%
  group_by(TransectID, Block) %>%
  arrange(Transect_position, .by_group = TRUE)

pepe.annual %>%
  filter(!is.na(Transect_position)) %>%
  group_by(TransectID) %>%
  summarise(n = n())

pepe.annual %>%
  filter(TransectID == 2)

## Transplants
## Assign random number to each location
DataSheets <- pepe.annual %>%
  # group_by(TransectID) %>%
  mutate(RandPlant = if_else(is.na(RandPlant), sample(1:100, length(is.na(RandPlant)), replace = FALSE), RandPlant),
         Season = if_else(is.na(Season), "Fall", Season))

pepe.annual.foo <- pepe.annual
pepe.annual.foo$RandPlant[is.na(pepe.annual.foo$RandPlant)] <- sample(1:100, 100, replace = FALSE)

table(pepe.annual.foo$RandPlant, pepe.annual.foo$TransectID)

pepe.annual.foo %>%
  group_by(TransectID) %>%
  arrange(Transect_position, .by_group = TRUE) %>%
  filter(is.na(Date) ) %>%
  write.csv(  "C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/2023Fall_Penstemon-penlandii.csv")
  
ExistingPlants <- pepe.annual %>%
  group_by(TransectID) %>%
  arrange(Transect_position, Y, X, .by_group = TRUE) %>%
  select(c(TransectID:PlantID,X,Y)) %>%
  filter(is.na(Transect_position))

ExistingPlants %>%
  group_by(TransectID) %>%
  arrange(Y,X, .by_group = TRUE) %>%
  write.csv(  "C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/2023ExistTrans_Penstemon-penlandii.csv")


pepe.20230626 <- pepe.20230626 %>%
  mutate(Date = as.Date("2023-06-26")) %>%
  inner_join(transBlock) %>%
  arrange(Block, TransectID)

pepe.spring <- pepe.spring %>%
  bind_rows(pepe.20230626)

pepe.3surv <- pepe.springfall %>%
  dplyr::select(c(TransectID, Season, Transect_position, xy, 
                  PlantID:Y, Date)) %>%
  inner_join(transBlock) %>%
  arrange(Block, TransectID) %>%
  mutate(Date = as.Date("2023/10/17"))

names(pepe.3surv) <- names(pepe.spring)
pepe.3surv <- pepe.3surv %>%
  bind_rows(pepe.spring)

```



```{r}
transplants <- pepe.all.2023 %>%
  left_join(pepe.transects, by = c("TransectID","Block")) %>%
  filter(!is.na(Transect_position)) %>%
  arrange(TransectID, Y, X)


## Spring transplants
## plot 4 all died, plot 6 are in twice - corrected in csv
transplants %>%
  filter(Season == "Spring") %>%
  filter(Date == "20231016") %>%
  filter(Height.cm > 0) %>%
  group_by(TransectID) %>%
  summarise(surv = n()) %>%
  full_join(pepe.transects[pepe.transects$Type_CT_TT == "TT",], by = "TransectID") %>%
  arrange(TransectID) %>%
  replace(is.na(.), 0) %>%
  group_by(TransectID, surv) %>%
  mutate(mu = median(dbinom(surv, 5, seq(0,1, by = 0.0001))),
         lower = HDInterval::hdi(dbinom(surv, 5, seq(0,1, by = 0.0001)), credMass = 0.95)[1],
         upper = HDInterval::hdi(dbinom(surv, 5, seq(0,1, by = 0.0001)), credMass = 0.95)[2]) %>%
  ggplot(  aes(as.factor(TransectID), mu)) +
    geom_point() +
    geom_errorbar(aes(ymin = lower, ymax = upper))+
    theme_bw() +
    xlab("Transect")+
    ylab("Survival rate") +
    scale_y_continuous(labels = scales::percent)

transplants %>%
  filter(Season == "Spring") %>%
  filter(Date == "20231016") %>%
  filter(Height.cm > 0) %>%
  group_by(TransectID) %>%
  summarise(surv = n()) %>%
  full_join(pepe.transects[pepe.transects$Type_CT_TT == "TT",], by = "TransectID") %>%
  arrange(TransectID) %>%
  replace(is.na(.), 0) %>%
  ggplot(  aes(as.factor(TransectID), surv/5)) +
    geom_point() +
    theme_bw() +
    xlab("Transect")+
    ylab("Survival rate") +
    scale_y_continuous(labels = scales::percent)

transplant_surv <- transplants %>%
  filter(Season == "Spring") %>%
  filter(Date == "20231016") %>%
  filter(Height.cm > 0) %>%
  group_by(TransectID) %>%
  summarise(surv = n()) %>%
  full_join(pepe.transects[pepe.transects$Type_CT_TT == "TT",], by = "TransectID") %>%
  arrange(TransectID) %>%
  replace(is.na(.), 0) %>%
  group_by(TransectID, surv) %>%
  mutate(mu = median(dbinom(surv, 5, seq(0,1, by = 0.0001))),
         lower = HDInterval::hdi(dbinom(surv, 5, seq(0,1, by = 0.0001)), credMass = 0.88)[1],
         upper = HDInterval::hdi(dbinom(surv, 5, seq(0,1, by = 0.0001)), credMass = 0.88)[2],
         percsurv = surv/5)

mean(transplant_surv$percsurv)

transplant_binom <- transplants %>%
  filter(Season == "Spring") %>%
  filter(Date == "20231016") %>%
  mutate(surv = case_when(Height.cm > 0 ~ 1,
                          Height.cm == 0 ~ 0)) 




## Density of surrounding plants in plot
existing <- pepe.all.2023 %>%
  left_join(pepe.transects, by = c("TransectID","Block")) %>%
  filter(is.na(Transect_position)) %>%
  group_by(TransectID, Block) %>%
  summarise(dens = n()/5) %>% ##sq meters
  arrange(TransectID)

trans_glm <- transplant_binom %>%
  left_join(existing, by = c("TransectID","Block")) %>%
  replace(is.na(.), 0)

lm1 <- glm(surv ~ dens, family = "binomial" , trans_glm)
summary(lm1)

#### This one!
ggsave(file = "C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/Figures/SpTransSurvxdensity.jpg", 
transplant_binom %>%
  left_join(existing, by = c("TransectID","Block")) %>%
  replace(is.na(.), 0) %>%
  ggplot(  aes(dens, surv)) +
    geom_point(pch = "|")+
    stat_smooth(method="glm",  method.args = list(family = "binomial"), se=TRUE)+
    theme_bw()+
    ylab("Spring transplant survival rate (May to Oct 2023)")+
    xlab("Exisiting density (Oct 2023)")
    , width=150, height=100,units='mm', dpi=300)  

existing %>%
  full_join(pepe.transects, by = c("TransectID","Block")) %>%
  filter(Type_CT_TT == "TT") %>%
  replace(is.na(.), 0) %>%
  arrange(TransectID) %>%
  left_join(transplant_surv) %>%
  select(TransectID:dens,OriginTagNum:End5mTagNum, surv, percsurv) %>%
  print(n=33)

## Because many died, need spring and 20230521 
transplants %>%
  filter(Height.cm == 0)
transplants <- transplants %>%
  mutate(Season = factor(Season, levels = c("Spring","Fall")))

ggsave(file = "C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/Figures/HeightWidthTransplantsXSeason.jpg", 
(transplants %>%
  filter(Season == "Spring" & Date == "20230531" | Season == "Fall") %>%
  left_join(pepe.transects[pepe.transects$Type_CT_TT == "TT",], by = c("TransectID","Block")) %>%
  arrange(TransectID) %>%
  ggplot(  aes(as.factor(TransectID), Height.cm, color = as.factor(Repro))) +
    geom_boxplot(outlier.shape = NA)+
    geom_point(position = position_dodge(width = 0.8), pch=16) +
    theme_bw() +
    facet_wrap(~Season,nrow=2)+
    xlab("Transect")+
    ylab("Height (cm)")+
    scale_color_manual("Stage", values = c("darkgreen","violet"), labels = c("Vegetative","Reproductive"))
) +
(transplants %>%
  filter(Season == "Spring" & Date == "20230531" | Season == "Fall") %>%
  left_join(pepe.transects[pepe.transects$Type_CT_TT == "TT",], by = c("TransectID","Block")) %>%
  arrange(TransectID) %>%
  ggplot(  aes(as.factor(TransectID), BasalWidth.cm, color = as.factor(Repro))) +
    geom_boxplot(outlier.shape = NA)+
    geom_point(position = position_dodge(width = 0.8), pch=16) +
    theme_bw() +
    facet_wrap(~Season,nrow=2)+
    xlab("Transect")+
    ylab("Width (cm)")+
    scale_color_manual("Stage", values = c("darkgreen","violet"), labels = c("Vegetative","Reproductive"))
 )

    , width=250, height=110,units='mm', dpi=300)  

transplants %>%
  filter(Season == "Spring" & Date == "20230531" | Season == "Fall") %>%
  select(X:Y,TransectID:Comments.x) %>%
  write.csv(., file = "C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/Figures/Transplants.csv")

transplants %>%
  filter(Season == "Spring" & Date == "20230531" | Season == "Fall") %>%
  select(X:Y,TransectID:Comments.x) %>%
  group_by(Season) %>%
  summarise(meanHeight = mean(Height.cm),
            meanWidth = mean(BasalWidth.cm),
            percRepro = sum(Repro)/100,
            meanFrFl = sum(NumRep))

pepe.all.2023 %>%
  filter(is.na(Transect_position)) %>%
  left_join(pepe.transects[pepe.transects$Type_CT_TT == "TT",], by = c("TransectID","Block")) %>%
  select(X:Y,TransectID:Block,Height.cm:Comments.x) %>%
  write.csv(., file = "C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/Figures/Existing.csv")

```



Confidence intervals on abundance from mark-recaptures
```{r}

library(recapr)

## n1 = individuals captured and marked in first sample
n1 <- pepe.all.2023 %>%
  left_join(pepe.transects, by = c("TransectID","Block")) %>%
  filter(is.na(Transect_position)) %>%
  filter(Type_CT_TT == "TT") %>%
  filter(Height.cm >0) %>%
  group_by(Date, TransectID) %>%
  summarise(n1 = n()) %>%
  arrange(TransectID, Date) %>%
  filter(Date == "20230531") %>%
  filter(TransectID != 18) 

n2 <- pepe.all.2023 %>%
  filter(is.na(Transect_position)) %>%
  left_join(pepe.transects, by = c("TransectID","Block")) %>%
  filter(Type_CT_TT == "TT") %>%
  filter(Height.cm >0) %>%
  group_by(Date, TransectID) %>%
  summarise(n2 = n()) %>%
  arrange(TransectID, Date) %>%
  filter(Date == "20231016") %>%
  filter(TransectID != 18)

#
m2 <- pepe.all.2023 %>%
  filter(is.na(Transect_position)) %>%
  left_join(pepe.transects, by = c("TransectID","Block")) %>%
  filter(Type_CT_TT == "TT") %>%
  filter(Height.cm >0) %>%
  group_by(PlantID, TransectID, Block) %>%
  summarize(n = n()) %>%
  filter(n == 2) %>%
  group_by(TransectID) %>%
  summarise(m2 = n())

recapt <- m2 %>%
  # select(TransectID:n1) %>%
  left_join(n1, by = "TransectID") %>%
  left_join(n2, by = "TransectID")

# Abundance
NChapman(n1 = recapt$n1, n2 = recapt$n2, m2 = recapt$m2)

for(i in 1:7){
 print(ciChapman(n1 = recapt$n1[i], n2 = recapt$n2[i], m2 = recapt$m2[i]) )
}


```


Detection among transplant plots and control plots (need to add that data)     
In manuscripts, cite this appendix for RMark and in describing it make sure to say something like “we used the R (R Development Core Team 2007) package RMark (Laake 2013) to construct models for program MARK (White and Burnham 1999).” Use citation("RMark") in R to get the proper citation for R.*

```{r}

PlantIDprobs <- c(309,1040)
pepe.all.2023 %>%
  filter(PlantID %in% PlantIDprobs)

## make capture history with early and late, Cannot include the transplanted individuals  
## Make a design matrix of groups, columns for each TransectID, 0 or 1 if in that Transect
pepe.dm <- pepe.all.2023 %>%
  filter(TransectID %in% recapt$TransectID) %>%
  filter(is.na(Transect_position)) %>%
  distinct(Date, TransectID, Block, PlantID, .keep_all = TRUE) %>%
  mutate(mark = if_else(Height.cm > 0, "1", "0")) %>%
  select(c(TransectID:Block, Date, mark, PlantID)) %>%
  mutate(Date = as.numeric(as.factor(Date))) %>%
  tidyr::complete(nesting(TransectID, Block, PlantID), Date, fill = list(mark = "0")) %>%
  arrange(Date) %>%
  pivot_wider(names_from = Date, values_from = mark, values_fill = "0", names_prefix = "Survey") %>%
  tidyr::unite(ch, Survey1:Survey2, sep="") %>%
  mutate(TransectID = as.factor(TransectID)) %>%
  filter(ch != "00")

## separate column for each factor, no intercept
dmTrans <- model.matrix(~ -1 + TransectID, pepe.dm)

pepe.dm %>%
  ungroup() %>%
  dplyr::select(c(ch)) %>%
  bind_cols(dmTrans) %>%
  dplyr::select(c(ch:TransectID40)) %>%
  mutate(TransectID40 = paste(TransectID40, ";", sep = "")) %>%
  write.table( file = "C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/2023ExistPePe.inp", 
               sep = " ", col.names = FALSE, row.names = FALSE)


transdf <- pepe.dm %>%
  ungroup() %>%
  dplyr::distinct(TransectID)

pepeRDTrans.inp <- convert.inp("C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/2023ExistPePe.inp",
                               group.df = transdf,
                               covariates = NULL,
                               use.comments = FALSE)

pepe.cjs <- process.data(pepeRDTrans.inp, model = "Closed", groups = "TransectID")
pepe.ddl <- make.design.data(pepe.cjs)

run.models <- function() {
  c.dot = list(formula =~ 1)
  c.plot = list(formula = ~TransectID)
  p.dot = list(formula =~1)
  p.plot = list(formula = ~TransectID)
  f0.dot = list(formula = ~1)
  f0.plot = list(formula = ~TransectID)
  
  mod.list <- create.model.list("Closed")
  
  results.out <- mark.wrapper(mod.list, data = pepe.cjs, ddl = pepe.ddl, output = FALSE)
  return(results.out)
}

pepe.results <- run.models()

## Model-selection table
pepe.results

pepe.results$p.plot.c.plot.f0.plot$results$derived %>%
  bind_cols(transdf)
pepe.results$p.dot.c.dot.f0.dot$results$derived %>%
  bind_cols(transdf)

pepe.results$p.dot.c.dot.f0.dot$results$real 



bestmod <- pepe.results$p.plot.c.plot.f0.dot$results$real
bestmod %>%
  mutate(rnam = row.names(.)) %>%
  separate(rnam, into = c("Parameter","Group","Time")) %>%
ggplot(  aes(x = Parameter, y = estimate, color = Parameter, group = Parameter))+
  geom_jitter()+
  geom_boxplot()+
  theme_bw()

bestmod %>%
  mutate(rnam = row.names(.)) %>%
  separate(rnam, into = c("Parameter","Group","Time")) %>%
  filter(Parameter == "f0")

textp.c.f0 <- bestmod %>%
  mutate(rnam = row.names(.)) %>%
  separate(rnam, into = c("Parameter","Group","Time")) %>%
  mutate(Group = gsub( "g", "",Group))

```



```{r}

fig6 <- pepe.all.2023 %>%
  filter(is.na(Transect_position)) %>%
  group_by(TransectID, Block) %>%
  summarise(Abundance = n()) %>%
  full_join(pepe.transects, by = c("TransectID","Block")) %>%
  # filter(Abundance > 0) %>%
  ggplot(  aes(TransectID, Abundance, fill = Type_CT_TT))+
    geom_bar(stat = "identity")+
    theme_bw()+
    geom_text(aes(label = round(Abundance/5,2)),vjust = 0.2,hjust = 0,
              size = 3, angle = 90) +
    xlab("Transects")+
    scale_fill_manual("Transect", values = c("darkgreen","grey"), labels = c("Control","Transplant"))
# +
#   geom_point(data = textp.c.f0, aes(Group, estimate*100, color = Parameter))

pepe.all.2023 %>%
  filter(is.na(Transect_position)) %>%
  group_by(TransectID, Block) %>%
  summarise(Abundance = n()) %>%
  full_join(pepe.transects, by = c("TransectID","Block")) %>%
  mutate(Type_CT_TT = as.factor(Type_CT_TT)) %>%
  replace(is.na(.), 0) %>%
  group_by(Type_CT_TT) %>%
  summarise(Density = round(mean(Abundance/5),2)) 
  
ggsave(file = "C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/Figures/exisitingpertreatment.jpg", fig6, width=175, height=100,units='mm', dpi=300)  
```





What percent are flowering in 2023-06-26 among the transplanted individuals?
```{r}
percFl <- pepe.3surv %>%
  mutate(Trans01 = if_else(!is.na(Transect_position), 1, 0)) %>%
  filter(Trans01 == 1) %>%
  group_by(TransectID, Date) %>%
  replace(is.na(.), 0) %>%
  summarise(PercFl = sum(Repro)/5)

pepe.3surv %>%
  filter(TransectID %in% c(30,32)) %>%
  replace(is.na(.), 0)

percFl %>%
  print(n=60)

ggplot(percFl, aes(Date, PercFl, color = as.factor(TransectID))) +
  geom_jitter(height = 0.0, width = 0.5) +
  # geom_line() +
  theme_bw()
  

## On average, percent reproduction increased by 17% (0% to 60% increase)
percFl %>%
  group_by(TransectID) %>%
  summarise(ChangePercFl = PercFl[2]-PercFl[1],
            Chang2 = PercFl[3]-PercFl[2]) %>%
  ungroup() %>%
  summarise(meanChange1 = mean(ChangePercFl),
            meanChange2 = mean(Chang2))


pepe.3surv %>%
  replace(is.na(.), 0) %>%
  mutate(Trans01 = if_else(!is.na(Transect_position), 1, 0)) %>%
  filter(Trans01 == 1) %>%
  group_by(TransectID, Date) %>%
  summarise(PercFl = sum(Repro)/5) %>%
  ungroup() %>%
  group_by(Date) %>%
  summarise(meanFl = mean(PercFl),
            minFl = min(PercFl),
            maxFl = max(PercFl))

## Survival from May to June to October
survSpring <- pepe.3surv %>%
  filter(PlantID < 6) %>%
  group_by(TransectID, Date) %>%
  summarise(Surv = length(Height.cm[Height.cm > 0])/5)

survSpring %>%
  ggplot(  aes(Date, Surv, color = as.factor(TransectID)))+
  geom_jitter(height = 0, width = 0.1)+
  geom_line() +
  theme_bw()

survSpring %>%
  group_by(Date) %>%
  # filter(Date == as.Date("2023-06-26")) %>%
  summarise(meanSurv = mean(Surv),
            minSurv = min(Surv),
            maxSurv = max(Surv))
```



## Root Shoot (only 2023)
```{r}
pepe.rootshoot <- read.csv("C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/2023_Penstemon-penlandii_rootshoot.csv")
```

```{r}

pepe.rootshoot$Season <- factor(pepe.rootshoot$Season, levels = c("Spring","Fall"))

# Not significant
ggsave(file = "C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/Figures/rootshootFig8.jpg", 

ggplot(pepe.rootshoot, aes( NumTertiaryRoots,Height, color = Trowel.BulbPlant
                            #, linetype = Season
                           )) +
  geom_point() +
  stat_smooth(method = "lm") +
  theme_bw()+
  xlab("Number of root hairs") +
  facet_wrap(~Season)+
  scale_color_manual("Method", values = c("grey60","black")) + 
  theme(legend.position="none") +
  
ggplot(pepe.rootshoot, aes( NumTertiaryRoots,Width, color = Trowel.BulbPlant
                            # , linetype = Season
                            )) +
  geom_point() +
  stat_smooth(method = "lm") +
  theme_bw()+
  xlab("Number of root hairs") +
  facet_wrap(~Season)+
  scale_color_manual("Method", values = c("grey60","black"))
    , width=150*2, height=100,units='mm', dpi=300)  


ggsave(file = "C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/Figures/rootshootWidthFig9.jpg", 

ggplot(pepe.rootshoot, aes( NumTertiaryRoots,Width, color = Trowel.BulbPlant
                            # ,linetype = Season
                            )) +
  geom_point() +
  stat_smooth(method = "lm") +
  theme_bw()+
  facet_wrap(~Season)+
  scale_color_manual("Method", values = c("grey60","black"))
    , width=150, height=100,units='mm', dpi=300)  

ggplot(pepe.rootshoot, aes(Taproot.diameter,Height,  color = Trowel.BulbPlant)) +
  geom_point() +
  stat_smooth(method = "lm") +
  theme_bw()+
  facet_wrap(~Season)

# Not significant
ggplot(pepe.rootshoot, aes(Num.lateral, Height, color = Trowel.BulbPlant)) +
  geom_point() +
  stat_smooth(method = "lm") +
  theme_bw()+
  facet_wrap(~Season)

## There are not more lateral roots as the taproot lengthens but there are more lateral roots later (the bulb planter wouldn't rip them off compared to the trowel)
ggplot(pepe.rootshoot, aes(Taprootlength, Num.lateral, color = Trowel.BulbPlant)) +
  geom_point() +
  stat_smooth(method = "lm") +
  theme_bw()+
  facet_wrap(~Season, scales = "free")

ggplot(pepe.rootshoot, aes(Num.lateral, Height)) +
  geom_point() +
  stat_smooth(method = "lm") +
  theme_bw()+
  facet_wrap(~Season)



## since bulb is later, number of lateral roots given taproot length which should also differ by technique
ggplot(pepe.rootshoot, aes(Num.lateral/Taprootlength, Height, color = Trowel.BulbPlant)) +
  geom_point() +
  stat_smooth(method = "lm") +
  theme_bw()+
  facet_wrap(~Season)

ggplot(pepe.rootshoot, aes(Taprootlength, Height, color = Trowel.BulbPlant)) +
  geom_point() +
  stat_smooth(method = "lm") +
  theme_bw()

pepe.rootshoot %>%
  # filter(Length.longest < 20) %>%
ggplot(  aes(Height, Length.longest, color = Trowel.BulbPlant)) +
  geom_point() +
  stat_smooth(method = "lm") +
  theme_bw()

# The number of root hairs does not depend on the length of the longest root
pepe.rootshoot %>%
  # filter(Length.longest < 20) %>%
ggplot(  aes(Length.longest, NumTertiaryRoots, color = Trowel.BulbPlant)) +
  stat_smooth(method = "lm") +
  geom_point() +
  theme_bw()+
  xlab("Length of longest root (cm)")+
  ylab("Root hairs")+
  scale_color_manual("Method", values = c("grey30","violet"))

# The number of root hairs depends on the number of lateral roots
pepe.rootshoot %>%
  # filter(Length.longest < 20) %>%
ggplot(  aes(Num.lateral, NumTertiaryRoots, color = Trowel.BulbPlant)) +
  stat_smooth(method = "lm") +
  geom_point() +
  theme_bw()+
  xlab("Lateral roots")+
  ylab("Root hairs")+
  scale_color_manual("Method", values = c("grey30","violet"))

# Number of lateral roots does not depend on length of longest root for either method
pepe.rootshoot %>%
  # filter(Length.longest < 20) %>%
ggplot(  aes(Length.longest, Num.lateral, color = Trowel.BulbPlant)) +
  stat_smooth(method = "lm") +
  geom_point() +
  theme_bw()+
  # xlab("Length of longest root (cm)")+
  # ylab("Root hairs")+
  scale_color_manual("Method", values = c("grey30","violet"))

# Number of root hairs/length of longest root larger for bulb (later) and trowel (two weeks earlier) 
pepe.rootshoot %>%
ggplot(  aes(NumTertiaryRoots/Length.longest, Trowel.BulbPlant)) +
  geom_boxplot()

## The number of lateral roots does not correlate with height for the bulb planter method (and later )
pepe.rootshoot %>%
ggplot(  aes(Height, Num.lateral, color = Trowel.BulbPlant)) +
  stat_smooth(method = "lm") +
  geom_point() +
  theme_bw()+
  # xlab("Length of longest root (cm)")+
  # ylab("Root hairs")+
  scale_color_manual("Method", values = c("grey30","violet"))

pepe.rootshoot %>%
ggplot(  aes(Height, Num.lateral)) +
  stat_smooth(method = "lm") +
  geom_point() +
  theme_bw()
  # xlab("Length of longest root (cm)")+
  # ylab("Root hairs")
  
  
pepe.rootshoot %>%
  # filter(Length.longest < 20) %>%
ggplot(  aes( Height,NumTertiaryRoots, color = Trowel.BulbPlant)) +
  stat_smooth(method = "lm") +
  geom_point() +
  theme_bw()+
  ylab("Root hairs")+
  xlab("Height (cm)")+
  scale_color_manual("Method", values = c("grey30","violet"))

pepe.rootshoot %>%
  # filter(Length.longest < 20) %>%
ggplot(  aes(Length.longest, Height, color = Trowel.BulbPlant)) +
  stat_smooth(method = "lm") +
  geom_point() +
  theme_bw()+
  xlab("Length of longest root (cm)")+
  ylab("Height (cm)")+
  scale_color_manual("Method", values = c("grey30","violet"))

ggplot(pepe.rootshoot, aes(Height, fill = Trowel.BulbPlant))+
  geom_density(alpha = 0.5)

ggplot(pepe.rootshoot, aes(Width, fill = Trowel.BulbPlant))+
  geom_density(alpha = 0.5)

## length of longest lateral root by timing/method
ggplot(pepe.rootshoot, aes(Length.longest, fill = Trowel.BulbPlant))+
  geom_boxplot(alpha = 0.5)

```


```{r}

panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...) {
    usr <- par("usr")
    on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    Cor <- abs(cor(x, y)) # Remove abs function if desired
    txt <- paste0(prefix, format(c(Cor, 0.123456789), digits = digits)[1])
    if(missing(cex.cor)) {
        cex.cor <- 0.4 / strwidth(txt)
    }
    text(0.5, 0.5, txt,
         cex = 1 + cex.cor * Cor) # Resize the text by level of correlation
}

panel.hist <- function(x, ...)
{
    usr <- par("usr")
    par(usr = c(usr[1:2], 0, 1.5) )
    h <- hist(x, plot = FALSE)
    breaks <- h$breaks; nB <- length(breaks)
    y <- h$counts; y <- y/max(y)
    rect(breaks[-nB], 0, breaks[-1], y, col = "grey", ...)
}
    
cor(pepe.rootshoot[,4:13])

reg <- function(x, y, col) abline(lm(y~x), col=col) 

panel.lm =  function (x, y, col = par("col"), bg = NA, pch = par("pch"), 
    cex = 1, col.smooth = "black", span = 2/3, iter = 3, ...)  {
    points(x, y, pch = pch, col = col, bg = bg, cex = cex)
    ok <- is.finite(x) & is.finite(y)
    if (any(ok)) reg(x[ok], y[ok], col.smooth)
}


```

levels are Trowel.Spring Bulb.Spring   Trowel.Fall   Bulb.Fall
```{r}
jpeg("C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/Root2Shootcorr.jpg",
     width=200, height=200,units='mm', res=300)


  # dplyr::select(c(Height, Width, Taprootlength, Taproot.diameter, Num.lateral, Length.longest, NumTertiaryRoots)) %>%
pairs(pepe.rootshoot[,c(4:5,8:12)], # pepe.rootshoot[pepe.rootshoot$Season == "Spring",c(4:5,8:12)],
      bg = c("blue","skyblue","darkred","pink")[unique(interaction(as.factor(pepe.rootshoot$Trowel.BulbPlant),
                                                            as.factor(pepe.rootshoot$Season)))],
      # bg = c("blue","darkred")[unique(pepe.rootshoot$Trowel.BulbPlant)],
      pch = 21,
      panel = panel.lm,
      lower.panel = panel.cor,
      diag.panel = panel.hist)

dev.off()
```


```{r}
# library(GGally)
# library(patchwork)

p1 <-pepe.rootshoot %>%
  filter(Season == "Spring") %>%
  rename(TapLn = Taprootlength, TapDiam = Taproot.diameter, NumLat = Num.lateral,
         LnLong = Length.longest, NumTert = NumTertiaryRoots) %>%
ggpairs(columns = c(4,8,10:12),
        aes(color = Trowel.BulbPlant, alpha = 0.25),
        upper = list(continuous = wrap("cor", size = 3)),
        lower = list(continuous = "smooth")) +
  ggtitle("a)")+
  theme_bw()

p2 <- pepe.rootshoot %>%
  filter(Season == "Fall") %>% 
  rename(TapLn = Taprootlength, TapDiam = Taproot.diameter, NumLat = Num.lateral,
         LnLong = Length.longest, NumTert = NumTertiaryRoots) %>%
ggpairs(columns = c(4,8,10:12),
        aes(color = Trowel.BulbPlant, alpha = 0.25),
        upper = list(continuous = wrap("cor", size = 3)),
        lower = list(continuous = "smooth"))+
  ggtitle("b)")+
  theme_bw()

ggsave(file = "C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/Figures/BulbVTrowelSpring.jpg", p1, width=150, height=150,units='mm', dpi=300)

ggsave(file = "C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/Figures/BulbVTrowelFall.jpg", p2, width=150, height=150,units='mm', dpi=300) 
```




TOMST data loggers   
```{r}
library(myClim)

datalogger1 <- mc_read_files(c("C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_lolly_TOMSTdatalogger/data_95135950_2023_10_17_0.csv", "C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_lolly_TOMSTdatalogger/data_95135951_2023_10_17_0.csv","C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_lolly_TOMSTdatalogger/data_95135952_2023_10_17_0.csv"),
                             dataformat_name = "TOMST")

files_csv <- system.file("extdata", "files_table.csv", package = "myClim")
localities_csv <- system.file("extdata", "localities_table.csv", package = "myClim")
tomst_data <- mc_read_data(files_csv, localities_csv)

  
  # mc_prep_clean(data = )
```

