---
title: "Pepe_transplant"
author: "Michelle DePrenger-Levin"
date: "2023-08-29"
output: html_document
---

---
```{r}

rm(list=ls()) 
library(R2jags)
library(runjags)
library(mcmcplots)
library(boot)
library(dplyr)
library(tidyr)
library(ggplot2)

library(RMark)
```



```{r}
pepe.spring <- read.csv("C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/2023Spring_Penstemon-penlandii.csv")

pepe.fall <- read.csv("C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/20231016Fall_Penstemon-penlandii_dataentry_2.csv")

## need to confirm 
pepe.20230626 <- read.csv("C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/20230626Spring_Penstemon-penlandii.csv")

## Need control plot data
pepe.20231016_control <- read.csv("C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/20231016_ControlTransects_dataentry.csv")

## Data from Fall 2023-10-16 and 2320-10-17 
pepe.springfall <- read.csv("C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/20231016Spring_Penstemon-penlandii_fall.csv")

```

Format for datasheets
```{r}


transBlock <- data.frame(TransectID = 1:40, Block = rep(1:20, each = 2))

pepe.spring <- pepe.spring %>%
  mutate(Date = as.Date("2023-05-31")) %>%
  inner_join(transBlock) %>%
  arrange(Block, TransectID)

Transect_positions <- data.frame(Transect_position = 1:10, xy = c("0.25, 0.25",
                                                                  "0.75, 0.75",
                                                                  "0.25, 1.25",
                                                                  "0.75, 1.75",
                                                                  "0.25, 2.25",
                                                                  "0.75, 2.75",
                                                                  "0.25, 3.25",
                                                                  "0.75, 3.75",
                                                                  "0.25, 4.25",
                                                                  "0.75, 4.75"))
  


## When only spring
pepe.annual <- pepe.spring %>%
  mutate(RandPlant = PlantID) %>%
  complete(nesting(TransectID, Block), Transect_positions) %>%
  group_by(TransectID, Block) %>%
  arrange(Transect_position, .by_group = TRUE)

pepe.annual %>%
  filter(!is.na(Transect_position)) %>%
  group_by(TransectID) %>%
  summarise(n = n())

pepe.annual %>%
  filter(TransectID == 2)

## Transplants
## Assign random number to each location
DataSheets <- pepe.annual %>%
  # group_by(TransectID) %>%
  mutate(RandPlant = if_else(is.na(RandPlant), sample(1:100, length(is.na(RandPlant)), replace = FALSE), RandPlant),
         Season = if_else(is.na(Season), "Fall", Season))

pepe.annual.foo <- pepe.annual
pepe.annual.foo$RandPlant[is.na(pepe.annual.foo$RandPlant)] <- sample(1:100, 100, replace = FALSE)

table(pepe.annual.foo$RandPlant, pepe.annual.foo$TransectID)

pepe.annual.foo %>%
  group_by(TransectID) %>%
  arrange(Transect_position, .by_group = TRUE) %>%
  filter(is.na(Date) ) %>%
  write.csv(  "C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/2023Fall_Penstemon-penlandii.csv")
  
ExistingPlants <- pepe.annual %>%
  group_by(TransectID) %>%
  arrange(Transect_position, Y, X, .by_group = TRUE) %>%
  select(c(TransectID:PlantID,X,Y)) %>%
  filter(is.na(Transect_position))

ExistingPlants %>%
  group_by(TransectID) %>%
  arrange(Y,X, .by_group = TRUE) %>%
  write.csv(  "C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/2023ExistTrans_Penstemon-penlandii.csv")


pepe.20230626 <- pepe.20230626 %>%
  mutate(Date = as.Date("2023-06-26")) %>%
  inner_join(transBlock) %>%
  arrange(Block, TransectID)

pepe.spring <- pepe.spring %>%
  bind_rows(pepe.20230626)

pepe.3surv <- pepe.springfall %>%
  dplyr::select(c(TransectID, Season, Transect_position, xy, 
                  PlantID:Y, Date)) %>%
  inner_join(transBlock) %>%
  arrange(Block, TransectID) %>%
  mutate(Date = as.Date("2023/10/17"))

names(pepe.3surv) <- names(pepe.spring)
pepe.3surv <- pepe.3surv %>%
  bind_rows(pepe.spring)

```

Detection among transplant plots and control plots (need to add that data)     
In manuscripts, cite this appendix for RMark and in describing it make sure to say something like “we used the R (R Development Core Team 2007) package RMark (Laake 2013) to construct models for program MARK (White and Burnham 1999).” Use citation("RMark") in R to get the proper citation for R.*

```{r}

## All individuals for annual 
pepe.annual <- pepe.springfall %>%
  mutate(IndivID = 1:nrow(pepe.springfall)) %>%
  mutate(Date = as.Date("2023/10/17"))

pepe.spring23 <- pepe.annual %>%
  dplyr::select(c(IndivID, X, Y, TransectID:Season, Transect_position, Height.cm:Comments)) %>%
  mutate(Date = as.Date("2023-05-31"))

pepe.fall23 <- pepe.annual %>%
  dplyr::select(c(IndivID, X, Y, TransectID:Season, Transect_position, Height.cm.1:Comments.1)) %>%
  # mutate(across(Height.cm:NumRep), as.numeric) %>%
  mutate(Date = as.Date("2023-10-17"))

# pepe.fall23[,7:10] <- sapply(pepe.fall23[,7:10],as.numeric)

names(pepe.fall23) <- names(pepe.spring23)

pepe.2023 <- bind_rows(pepe.spring23, pepe.fall23)

## make capture history with early and late, Cannot include the transplanted individuals  
## Make a design matrix of groups, columns for each TransectID, 0 or 1 if in that Transect
pepe.dm <- pepe.2023 %>%
  filter(is.na(Transect_position)) %>%
  filter(!is.na(Height.cm)) %>%
  mutate(Trans01 = if_else(!is.na(Transect_position), 1, 0)) %>%
  mutate(mark = if_else(Height.cm > 0, 1, 0)) %>%
  dplyr::select(Date,mark,IndivID, Trans01, TransectID) %>%
  pivot_wider(names_from = Date, names_prefix = "Survey", values_from = mark, values_fill = 0) %>%
  unite(ch, `Survey2023-05-31`:`Survey2023-10-17`, sep="") %>%
  mutate(TransectID = as.factor(TransectID))

## separate column for each factor, no intercept
dmTrans <- model.matrix(~ -1 + TransectID, pepe.dm)
dmTrans <- dmTrans[,colSums(dmTrans) != 0]

## Phi and p
# dm <- cbind(dmTrans,matrix(0,nrow=nrow(pepe.dm),ncol=length(colnames(dmTrans))))
# dm <- rbind(dm,cbind(matrix(0,nrow=nrow(pepe.dm),ncol=length(colnames(dmTrans))),dmTrans))
# colnames(dm) <- c(paste("p:", colnames(dmTrans), sep=""),
#                   paste("Phi:", colnames(dmTrans), sep=""))

pepe.2023 %>%
  filter(is.na(Transect_position)) %>%
  filter(!is.na(Height.cm)) %>%
  ## Seems to have no data for 16, 22, or 24
  # filter(!(TransectID %in% c(16,22,24))) %>%
  filter(TransectID != 16) %>%
  mutate(Trans01 = if_else(!is.na(Transect_position), 1, 0)) %>%
  mutate(mark = if_else(Height.cm > 0, 1, 0)) %>%
  dplyr::select(Date,mark,IndivID, Trans01, TransectID) %>%
  pivot_wider(names_from = Date, names_prefix = "Survey", values_from = mark, values_fill = 0) %>%
  unite(ch, `Survey2023-05-31`:`Survey2023-10-17`, sep="") %>%
  mutate(Trans10 = if_else(TransectID == 10, 1, 0)) %>%
  mutate(Trans12 = if_else(TransectID == 12, 1, 0)) %>%
  mutate(Trans18 = if_else(TransectID == 18, 1, 0)) %>%
  # mutate(Trans16 = if_else(TransectID == 16, 1, 0)) %>%
  mutate(Trans20 = if_else(TransectID == 20, 1, 0)) %>%
  mutate(Trans22 = if_else(TransectID == 22, 1, 0)) %>%
  mutate(Trans24 = if_else(TransectID == 24, 1, 0)) %>%
  mutate(Trans26 = if_else(TransectID == 26, 1, 0)) %>%
  mutate(Trans32 = if_else(TransectID == 32, 1, 0)) %>%
  mutate(Trans34 = if_else(TransectID == 34, 1, 0)) %>%
  mutate(Trans36 = if_else(TransectID == 36, 1, 0)) %>%
  mutate(Trans38 = if_else(TransectID == 38, 1, 0)) %>%
  mutate(Trans40 = if_else(TransectID == 40, "1;", "0;")) %>%
  # bind_cols(dmTrans[,-1]) %>%
  # mutate(TransectID40 = "TransectID40;") %>%
  dplyr::select(c(ch:Trans40)) %>%
  write.table( file = "C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/2023ExistTrans_Penstemon-penlandii.inp", sep = " ", col.names = FALSE, row.names = FALSE)

transInclude <- pepe.2023 %>%
  filter(is.na(Transect_position)) %>%
  filter(!is.na(Height.cm)) %>%
  # filter(!(TransectID %in% c(16,22,24))) %>%
  filter(TransectID != 16) %>%
  mutate(mark = if_else(Height.cm > 0, 1, 0)) %>%
  dplyr::select(Date,mark,IndivID, TransectID) %>%
  pivot_wider(names_from = Date, names_prefix = "Survey", values_from = mark, values_fill = 0) %>%
  distinct(TransectID) %>%
  mutate(TransectID = as.factor(TransectID))

pepe.inp <- convert.inp("C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/2023ExistTrans_Penstemon-penlandii.inp", 
                        group.df = transInclude, covariates = NULL)


pepe.cjs <- process.data(pepe.inp, model = "Closed", groups = "TransectID")
pepe.ddl <- make.design.data(pepe.cjs)

pepe.ddl$p$group
pepe.ddl$c$group

run.models <- function() {
  c.dot = list(formula =~ 1)
  c.plot = list(formula = ~TransectID)
  p.dot = list(formula =~1)
  p.plot = list(formula = ~TransectID)
  f0.dot = list(formula = ~1)
  f0.plot = list(formula = ~TransectID)
  
  mod.list <- create.model.list("Closed")
  
  results.out <- mark.wrapper(mod.list, data = pepe.cjs, ddl = pepe.ddl, output = FALSE)
  return(results.out)
}

pepe.results <- run.models()

## Model-selection table
pepe.results

bestmod <- pepe.results$p.plot.c.plot.f0.dot$results$real

bestmod %>%
  mutate(rnam = row.names(.)) %>%
  separate(rnam, into = c("Parameter","Group","Time")) %>%
ggplot(  aes(x = Parameter, y = estimate, color = Parameter, group = Parameter))+
  geom_jitter()+
  geom_boxplot()+
  theme_bw()

29/nrow(pepe.2023)

textp.c.f0 <- bestmod %>%
  mutate(rnam = row.names(.)) %>%
  separate(rnam, into = c("Parameter","Group","Time")) %>%
  mutate(Group = gsub( "g", "",Group))

```

```{r}

fig6 <- pepe.2023 %>%
  filter(is.na(Transect_position)) %>%
  group_by(TransectID) %>%
  summarise(Abundance = n()) %>%
  filter(Abundance > 0) %>%
  ggplot(  aes(TransectID, Abundance))+
  geom_bar(stat = "identity")+
  theme_bw()+
  geom_text(aes(label = round(Abundance/5,2)), vjust = -0.2) +
  xlab("Transects")
# +
#   geom_point(data = textp.c.f0, aes(Group, estimate*100, color = Parameter))
  
ggsave(file = "C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/Figures/exisitingpertreatment.jpg", fig6, width=150, height=125,units='mm', dpi=300)  
```




What percent are flowering in 2023-06-26 among the transplanted individuals?
```{r}
percFl <- pepe.3surv %>%
  mutate(Trans01 = if_else(!is.na(Transect_position), 1, 0)) %>%
  filter(Trans01 == 1) %>%
  group_by(TransectID, Date) %>%
  replace(is.na(.), 0) %>%
  summarise(PercFl = sum(Repro)/5)

pepe.3surv %>%
  filter(TransectID %in% c(30,32)) %>%
  replace(is.na(.), 0)

percFl %>%
  print(n=60)

ggplot(percFl, aes(Date, PercFl, color = as.factor(TransectID))) +
  geom_jitter(height = 0.0, width = 0.5) +
  # geom_line() +
  theme_bw()
  

## On average, percent reproduction increased by 17% (0% to 60% increase)
percFl %>%
  group_by(TransectID) %>%
  summarise(ChangePercFl = PercFl[2]-PercFl[1],
            Chang2 = PercFl[3]-PercFl[2]) %>%
  ungroup() %>%
  summarise(meanChange1 = mean(ChangePercFl),
            meanChange2 = mean(Chang2))


pepe.3surv %>%
  replace(is.na(.), 0) %>%
  mutate(Trans01 = if_else(!is.na(Transect_position), 1, 0)) %>%
  filter(Trans01 == 1) %>%
  group_by(TransectID, Date) %>%
  summarise(PercFl = sum(Repro)/5) %>%
  ungroup() %>%
  group_by(Date) %>%
  summarise(meanFl = mean(PercFl),
            minFl = min(PercFl),
            maxFl = max(PercFl))

## Survival from May to June to October
survSpring <- pepe.3surv %>%
  filter(PlantID < 6) %>%
  group_by(TransectID, Date) %>%
  summarise(Surv = length(Height.cm[Height.cm > 0])/5)

survSpring %>%
  ggplot(  aes(Date, Surv, color = as.factor(TransectID)))+
  geom_jitter(height = 0, width = 0.1)+
  geom_line() +
  theme_bw()

survSpring %>%
  group_by(Date) %>%
  # filter(Date == as.Date("2023-06-26")) %>%
  summarise(meanSurv = mean(Surv),
            minSurv = min(Surv),
            maxSurv = max(Surv))
```




```{r}
pepe.rootshoot <- read.csv("C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/2023_Penstemon-penlandii_rootshoot.csv")
```

```{r}

pepe.rootshoot$Season <- factor(pepe.rootshoot$Season, levels = c("Spring","Fall"))

# Not significant
ggplot(pepe.rootshoot, aes( NumTertiaryRoots,Height, color = Trowel.BulbPlant,
                            linetype = Season)) +
  geom_point() +
  stat_smooth(method = "lm") +
  theme_bw()+
  facet_wrap(~Season)

ggplot(pepe.rootshoot, aes(Taproot.diameter,Height,  color = Trowel.BulbPlant)) +
  geom_point() +
  stat_smooth(method = "lm") +
  theme_bw()+
  facet_wrap(~Season)

# Not significant
ggplot(pepe.rootshoot, aes(Num.lateral, Height, color = Trowel.BulbPlant)) +
  geom_point() +
  stat_smooth(method = "lm") +
  theme_bw()+
  facet_wrap(~Season)

## There are not more lateral roots as the taproot lengthens but there are more lateral roots later (the bulb planter wouldn't rip them off compared to the trowel)
ggplot(pepe.rootshoot, aes(Taprootlength, Num.lateral, color = Trowel.BulbPlant)) +
  geom_point() +
  stat_smooth(method = "lm") +
  theme_bw()+
  facet_wrap(~Season, scales = "free")

ggplot(pepe.rootshoot, aes(Num.lateral, Height)) +
  geom_point() +
  stat_smooth(method = "lm") +
  theme_bw()+
  facet_wrap(~Season)



## since bulb is later, number of lateral roots given taproot length which should also differ by technique
ggplot(pepe.rootshoot, aes(Num.lateral/Taprootlength, Height, color = Trowel.BulbPlant)) +
  geom_point() +
  stat_smooth(method = "lm") +
  theme_bw()+
  facet_wrap(~Season)

ggplot(pepe.rootshoot, aes(Taprootlength, Height, color = Trowel.BulbPlant)) +
  geom_point() +
  stat_smooth(method = "lm") +
  theme_bw()

pepe.rootshoot %>%
  # filter(Length.longest < 20) %>%
ggplot(  aes(Height, Length.longest, color = Trowel.BulbPlant)) +
  geom_point() +
  stat_smooth(method = "lm") +
  theme_bw()

# The number of root hairs does not depend on the length of the longest root
pepe.rootshoot %>%
  # filter(Length.longest < 20) %>%
ggplot(  aes(Length.longest, NumTertiaryRoots, color = Trowel.BulbPlant)) +
  stat_smooth(method = "lm") +
  geom_point() +
  theme_bw()+
  xlab("Length of longest root (cm)")+
  ylab("Root hairs")+
  scale_color_manual("Method", values = c("grey30","violet"))

# The number of root hairs depends on the number of lateral roots
pepe.rootshoot %>%
  # filter(Length.longest < 20) %>%
ggplot(  aes(Num.lateral, NumTertiaryRoots, color = Trowel.BulbPlant)) +
  stat_smooth(method = "lm") +
  geom_point() +
  theme_bw()+
  xlab("Lateral roots")+
  ylab("Root hairs")+
  scale_color_manual("Method", values = c("grey30","violet"))

# Number of lateral roots does not depend on length of longest root for either method
pepe.rootshoot %>%
  # filter(Length.longest < 20) %>%
ggplot(  aes(Length.longest, Num.lateral, color = Trowel.BulbPlant)) +
  stat_smooth(method = "lm") +
  geom_point() +
  theme_bw()+
  # xlab("Length of longest root (cm)")+
  # ylab("Root hairs")+
  scale_color_manual("Method", values = c("grey30","violet"))

# Number of root hairs/length of longest root larger for bulb (later) and trowel (two weeks earlier) 
pepe.rootshoot %>%
ggplot(  aes(NumTertiaryRoots/Length.longest, Trowel.BulbPlant)) +
  geom_boxplot()

## The number of lateral roots does not correlate with height for the bulb planter method (and later )
pepe.rootshoot %>%
ggplot(  aes(Height, Num.lateral, color = Trowel.BulbPlant)) +
  stat_smooth(method = "lm") +
  geom_point() +
  theme_bw()+
  # xlab("Length of longest root (cm)")+
  # ylab("Root hairs")+
  scale_color_manual("Method", values = c("grey30","violet"))

pepe.rootshoot %>%
ggplot(  aes(Height, Num.lateral)) +
  stat_smooth(method = "lm") +
  geom_point() +
  theme_bw()
  # xlab("Length of longest root (cm)")+
  # ylab("Root hairs")
  
  
pepe.rootshoot %>%
  # filter(Length.longest < 20) %>%
ggplot(  aes( Height,NumTertiaryRoots, color = Trowel.BulbPlant)) +
  stat_smooth(method = "lm") +
  geom_point() +
  theme_bw()+
  ylab("Root hairs")+
  xlab("Height (cm)")+
  scale_color_manual("Method", values = c("grey30","violet"))

pepe.rootshoot %>%
  # filter(Length.longest < 20) %>%
ggplot(  aes(Length.longest, Height, color = Trowel.BulbPlant)) +
  stat_smooth(method = "lm") +
  geom_point() +
  theme_bw()+
  xlab("Length of longest root (cm)")+
  ylab("Height (cm)")+
  scale_color_manual("Method", values = c("grey30","violet"))

ggplot(pepe.rootshoot, aes(Height, fill = Trowel.BulbPlant))+
  geom_density(alpha = 0.5)

ggplot(pepe.rootshoot, aes(Width, fill = Trowel.BulbPlant))+
  geom_density(alpha = 0.5)

## length of longest lateral root by timing/method
ggplot(pepe.rootshoot, aes(Length.longest, fill = Trowel.BulbPlant))+
  geom_boxplot(alpha = 0.5)

```


```{r}

panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...) {
    usr <- par("usr")
    on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    Cor <- abs(cor(x, y)) # Remove abs function if desired
    txt <- paste0(prefix, format(c(Cor, 0.123456789), digits = digits)[1])
    if(missing(cex.cor)) {
        cex.cor <- 0.4 / strwidth(txt)
    }
    text(0.5, 0.5, txt,
         cex = 1 + cex.cor * Cor) # Resize the text by level of correlation
}

panel.hist <- function(x, ...)
{
    usr <- par("usr")
    par(usr = c(usr[1:2], 0, 1.5) )
    h <- hist(x, plot = FALSE)
    breaks <- h$breaks; nB <- length(breaks)
    y <- h$counts; y <- y/max(y)
    rect(breaks[-nB], 0, breaks[-1], y, col = "grey", ...)
}
    
cor(pepe.rootshoot[,4:13])
```

```{r}
jpeg("C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_Transplant_2023-2026/Root2Shootcorr.jpg",
     width=200, height=200,units='mm', res=300)

pairs(pepe.rootshoot[,c(4:5,8:12)],
      bg = rainbow(2)[as.factor(pepe.rootshoot$Trowel.BulbPlant)],
      # col = rainbow(2)[as.factor(pepe.rootshoot$Trowel.BulbPlant)],
      pch = 21,
      lower.panel = panel.cor,
      diag.panel = panel.hist)
      # lower.panel = panel.smooth)

dev.off()
```



TOMST data loggers   
```{r}
library(myClim)

datalogger1 <- mc_read_files(c("C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_lolly_TOMSTdatalogger/data_95135950_2023_10_17_0.csv", "C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_lolly_TOMSTdatalogger/data_95135951_2023_10_17_0.csv","C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Penstemon-penlandii/Penstemon-penlandii_Projects/Tri-State_lolly_TOMSTdatalogger/data_95135952_2023_10_17_0.csv"),
                             dataformat_name = "TOMST")

files_csv <- system.file("extdata", "files_table.csv", package = "myClim")
localities_csv <- system.file("extdata", "localities_table.csv", package = "myClim")
tomst_data <- mc_read_data(files_csv, localities_csv)

  
  # mc_prep_clean(data = )
```

