---
title: "Penstemon penlandii transplant study"
author: "Michelle DePrenger-Levin"
date: "2023-02-06"
output:
  word_document: default
  html_document: default
  pdf_document: default
---
```{r}

rm(list=ls()) 
library(R2jags)
library(runjags)
library(mcmcplots)
library(boot)
```



Want to know spring or fall (make a dag)

1. Poisson for reproduction output
2.Simulate beta distribution for survival of stages [0,1]
3. binomial for being reproductive or not
both have parameters: 
normal for Beta_transplanttime impact of spring vs. fall
lognormal for Beta_size impact of size  

Put together for MPM and estimate the lambda (or an IPM) 
#########################################################################################################  
  simple model of survival (no random effects, no grouping)
  Penstemon penlandii model and data to test
  
```{r,"PePe"}
pepestring <- '
  data{
    # Likelihood:
        for(i in 1:N){
          y[i] ~ dbern(theta[i]) ## theta: The probability of survival
          theta[i] <- alpha + beta * x[i]  
        }
    }
  model{
  fake <- 0
  }
  '
# parameters for simulations 
N = 400 # nb of observations
x <- as.numeric(as.factor(rep(c("trSp","trFa","coSo","coDe"), each = 100))) # predictor
alpha <- 0.4
beta = 0.1 # offset of mean survival
# theta <- .5 # probability of survival (variance is theta(1-theta))
# parameters are treated as data for the simulation step
data<-list(N=N,x=x,alpha=alpha,beta=beta)

# Run JAGS but monitor the response variable instead of the parameters because we're checking if the model works 
out <- run.jags(pepestring, data = data,monitor=c("y"),sample=1, n.chains=1, summarise=FALSE)

Simulated <- coda::as.mcmc(out)
## spit out y for each data point
dat <- as.vector(Simulated)
```
# Now fit the model used to simulate the data that was generated
```{r}
# specify model in BUGS language
modelpepe <- 	
paste("	
model {
    # Likelihood:
        for(i in 1:N){
          y[i] ~ dbern(theta[i]) ## theta: The probability of survival
          theta[i] <- alpha + beta * x[i]  
        }
    # Priors:
    alpha ~ dbeta(1, 1) # intercept ## uniform, uninformative prior
    beta ~ dnorm(0, 0.01) # slope
  }
")
writeLines(modelpepe,"lin_reg_pepe.jags") ## written to  "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/Documents"
# setwd("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/Documents")

# data, dat from simulation
jags.data <- list(y = dat, N = length(dat), x = x)

# initial values
inits <- function(){list(alpha = rbeta(1,1,1), beta = rnorm(1))}  

# parameters monitored
parameters <- c("alpha", "beta")

# MCMC settings
ni <- 10000
nt <- 6
nb <- 5000
nc <- 2

# call JAGS from R
res <- jags(jags.data, inits = NULL, 
            # inits = inits, 
            parameters, "lin_reg_pepe.jags", n.chains = nc, n.thin = nt, 
            n.iter = ni, n.burnin = nb,
            working.directory = getwd()) 

print(res, digits = 3)

## Check convergence 
# trace plots
traplot(res,c("alpha", "beta"))
# posterior distributions
denplot(res,c("alpha", "beta"))
```





################################################################################################################################################################################################################  
  simple model of survival (grouping by treatment and control)
  Penstemon penlandii model and data to test
  
```{r,"PePeTC"}
modelpepe <- '	
data {
  for(i in 1:N) {
    y[i] ~ dbern(p[i])
    logit(p[i]) <- alpha[ ct[i] ]  ## one for each plot
  }
}

  model{
  fake <- 0
  }
  '


# parameters for simulations 
N = 400 # nb of observations
TC = 4
ct <- as.factor(rep(c("trSp","trFa","coSo","coDe"), each = 100)) # predictor, 4,3,2,1
meanAlpha <- mean(c(1.1,1,-1,-2)) ## because centered, looking for difference from mean
alpha <- c(1.1,1,-1,-2) # index:  coDe = 1, coSo =2,  trFa =3, trSp=4
inv.logit(alpha) ## 0.7502601 0.7310586 0.2689414 0.1192029 are the probabilities
# parameters are treated as data for the simulation step
data<-list(N=N,ct=ct, alpha=alpha)

# Run JAGS but monitor the response variable instead of the parameters because we're checking if the model works 
out <- run.jags(modelpepe, data = data,monitor=c("y"),sample=1, n.chains=1, summarise=FALSE)

Simulated <- coda::as.mcmc(out)
## spit out y for each data point
dat <- as.vector(Simulated)

```

 <https://mmeredith.net/blog/2017/Categories_in_JAGS.htm>     
# fit model center caegorical coefficients (subtract out the mean)
```{r}
# specify model in BUGS language
modelpepe <- 	
paste("	
model {
  for(i in 1:N) {
    y[i] ~ dbern(p[i])
    logit(p[i]) <- alpha[ ct[i] ]  ## one for each plot
  }

  # Priors
    for(i in 1:4) {
      Plottemp[i] ~ dunif(-5,5)   ## find the mean across plotTypes
    }
  alpha <- Plottemp - mean(Plottemp) ## resulting in change in centered Plot for each type
}")
writeLines(modelpepe,"centered_pepe.jags")

# data, dat from simulation
jags.data <- list(y = dat, ct = ct, N = length(dat))

# initial values
inits <- function(){list(alpha = rnorm(1))}  

# parameters monitored
parameters <- c("alpha")

# MCMC settings
ni <- 10000
nt <- 6
nb <- 5000
nc <- 2

# call JAGS from R
res <- jags(jags.data, inits = NULL, parameters, 
            "centered_pepe.jags", n.chains = nc, n.thin = nt, 
            n.iter = ni, n.burnin = nb,
            working.directory = getwd()) 

print(res, digits = 3)

## Check convergence 
# trace plots
traplot(res,c("alpha", "bPlot"))
# posterior distributions
denplot(res,c("alpha", "bPlot"))
```

#######################################################################################################



############################# Centered categorical, size ##############################################


    ## b0=alpha, bPlayer=bPlot, player=ct, bFluff=bSize, fluff=size
# fit model center caegorical coefficients (subtract out the mean)
```{r}  
modelpepeSize <- 	'
data {
  for(i in 1:N) {
    y[i] ~ dbern(p[i])
    logit(p[i]) <- bPlot[ct[i]] + bSize * size[i]
  }
}

  model{
  fake <- 0
  }
  '

# parameters for simulations 
N = 400 # nb of observations
TC = 4
ct <- as.numeric(as.factor(rep(c("trSp","trFa","coSo","coDe"), each = 100))) # predictor, 4,3,2,1
bPlot <- c(1.1,1,-1,-2) # index:  coDe = 1, coSo =2,  trFa =3, trSp=4
bSize <- 1
size <- rlnorm(N, log(5), 1)
# parameters are treated as data for the simulation step
data<-list(N=N,ct=ct, bPlot=bPlot, bSize = bSize, size = size)
# Run JAGS but monitor the response variable instead of the parameters because we're checking if the model works 
out <- run.jags(modelpepeSize, data = data,monitor=c("y"),sample=1, n.chains=1, summarise=FALSE)
Simulated <- coda::as.mcmc(out)
## spit out y for each data point
dat <- as.vector(Simulated)

# specify model in BUGS language
modelpepeSize <- 	
paste("	
model {
  for(i in 1:N) {
    y[i] ~ dbern(p[i])
    logit(p[i]) <- bPlot[ct[i]] + bSize * size[i]
    }

  # Priors
  alpha ~ dunif(-5,5) ## because it is centered 
    for(i in 1:4) {
      Plottemp[i] ~ dnorm(0, 0.01)   ## find the mean across plotTypes
    }
  bPlot <- Plottemp - mean(Plottemp) ## resulting in change in centered Plot for each type
  bSize ~ dunif(-5,5)
}")
writeLines(modelpepeSize,"centered_pepeSize.jags")

# data, dat from simulation
jags.data <- list(y = dat, ct = ct, N = length(dat), size = size)

# initial values
# inits <- function(){list(alpha = rbeta(1,1,1), beta = rnorm(1))}  

# parameters monitored
parameters <- c("bPlot", "bSize")

# MCMC settings
ni <- 10000
nt <- 6
nb <- 5000
nc <- 2

# call JAGS from R
res <- jags(jags.data, inits = NULL, parameters, 
            "centered_pepeSize.jags", n.chains = nc, n.thin = nt, 
            n.iter = ni, n.burnin = nb,
            working.directory = getwd()) 

print(res, digits = 3)

## Check convergence 
# trace plots
traplot(res,c( "bPlot","bSize"))
# posterior distributions
denplot(res,c( "bPlot","bSize"))
```

#######################################################################################################


############################# Centered categorical, size, centered stage ##############################################


    ## b0=alpha, bPlayer=bPlot, player=ct, bFluff=bSize, fluff=size
# fit model center caegorical coefficients (subtract out the mean)
```{r}  
# parameters for simulations 
N = 400 # nb of observations
TC = 4
ct <- as.factor(rep(c("trSp","trFa","coSo","coDe"), each = 100)) # predictor, 4,3,2,1
alpha <- mean(c(1.1,1,-1,-2)) ## because centered, looking for difference from mean
bPlot <- c(1.1,1,-1,-2) # index:  coDe = 1, coSo =2,  trFa =3, trSp=4
bSize <- 1
size <- rlnorm(N, log(5), 1)
bStage <- 1.1 ## slightly more likely to survive as a reproductive than juvenile (long-lived perennial) ???
## prob of rep depends on size, 0=veg, 1=rep
alphaVegRep <- 0
bVegRep <- 0.1 # slightly more likely to reproductive as get bigger
stage <- unlist(lapply(inv.logit(alphaVegRep + bVegRep*size), function(x) rbinom(n=1, size=1, prob=x)))  
# parameters are treated as data for the simulation step
dat <- unlist(lapply(inv.logit(alpha + bPlot[ct] + size*bSize + stage*bStage), function(x) rbinom(n = 1, size = 1, prob = x)))

data.frame(ct, size, stage)

# specify model in BUGS language
modelpepeSize <- 	
paste("	
model {
  # Likelihood
    for(i in 1:N) {
      y[i] ~ dbern(p[i])
      logit(p[i]) <- alpha + bPlot[ct[i]] + bSize * size[i] + bStage * stage[i]
      
      stage[i] ~ dbern(pS[i])
      logit(pS[i]) <- alphaVegRep + bVegRep * size[i]
    }
    
    # Stage, Hierarchical based on size, survival based on stage and size (or maybe confounded)
    # for(i in 1:N)
    # }

  # Priors
  alpha ~ dunif(-5,5) ## because it is centered 
    for(i in 1:4) {
      Plottemp[i] ~ dnorm(0, 0.01)   ## find the mean across plotTypes
    }
  bPlot <- Plottemp - mean(Plottemp) ## resulting in change in centered Plot for each type
  bSize ~ dunif(-5,5)
  bStage ~ dunif(-5,5)
  alphaVegRep ~ dbern(0.5)
  bVegRep ~ dunif(-5,5)
  
}")
writeLines(modelpepeSize,"centered_pepeSize.jags")

# data, dat from simulation
jags.data <- list(y = dat, ct = ct, N = length(dat), size = size)

# initial values
# inits <- function(){list(alpha = rbeta(1,1,1), beta = rnorm(1))}  

# parameters monitored
parameters <- c("bPlot", "alpha", "bSize", "bStage", "alphaVegRep","bVegRep")

# MCMC settings
ni <- 10000
nt <- 6
nb <- 5000
nc <- 2

# call JAGS from R
res <- jags(jags.data, inits = NULL, parameters, 
            "centered_pepeSize.jags", n.chains = nc, n.thin = nt, 
            n.iter = ni, n.burnin = nb,
            working.directory = getwd()) 

print(res, digits = 3)

## Check convergence 
# trace plots
traplot(res,c("alpha", "bPlot","bSize", "bStage", "alphaVegRep","bVegRep"))
# posterior distributions
denplot(res,c("alpha", "bPlot","bSize", "bStage", "alphaVegRep","bVegRep"))
```

#######################################################################################################


############################# Centered categorical, size, centered stage ##############################################


    ## b0=alpha, bPlayer=bPlot, player=ct, bFluff=bSize, fluff=size
# fit model center caegorical coefficients (subtract out the mean)
```{r}  
# parameters for simulations 
N = 400 # nb of observations
TC = 4
ct <- as.factor(rep(c("trSp","trFa","coSo","coDe"), each = 100)) # predictor, 4,3,2,1
alpha <- mean(c(1.1,1,-1,-2)) ## because centered, looking for difference from mean
bPlot <- c(1.1,1,-1,-2) # index:  coDe = 1, coSo =2,  trFa =3, trSp=4
bSize <- 1
size <- rlnorm(N, log(5), 1)
bStage <- 1.1 ## slightly more likely to survive as a reproductive than juvenile (long-lived perennial) ???
## prob of rep depends on size, 0=veg, 1=rep
alphaVegRep <- 0
bVegRep <- 0.1 # slightly more likely to reproductive as get bigger
stage <- unlist(lapply(inv.logit(alphaVegRep + bVegRep*size), function(x) rbinom(n=1, size=1, prob=x)))  
# parameters are treated as data for the simulation step
dat <- unlist(lapply(inv.logit(alpha + bPlot[ct] + size*bSize + stage*bStage), function(x) rbinom(n = 1, size = 1, prob = x)))

data.frame(ct, size, stage)

# specify model in BUGS language
modelpepeSize <- 	
paste("	
model {
  # Likelihood
    for(i in 1:N) {
      y[i] ~ dbern(p[i])
      logit(p[i]) <- alpha + bPlot[ct[i]] + bSize * size[i] + bStage * stage[i]
    }
    
    # Stage, Hierarchical based on size, survival based on stage and size (or maybe confounded)
    # for(i in 1:N)
    # }

  # Priors
  alpha ~ dunif(-5,5) ## because it is centered 
    for(i in 1:4) {
      Plottemp[i] ~ dnorm(0, 0.01)   ## find the mean across plotTypes
    }
  bPlot <- Plottemp - mean(Plottemp) ## resulting in change in centered Plot for each type
  bSize ~ dunif(-5,5)
  bStage ~ dunif(-5,5)
  
}")
writeLines(modelpepeSize,"centered_pepeSize.jags")

# data, dat from simulation
jags.data <- list(y = dat, ct = ct, N = length(dat), size = size, stage = stage)

# initial values
# inits <- function(){list(alpha = rbeta(1,1,1), beta = rnorm(1))}  

# parameters monitored
parameters <- c("bPlot", "alpha", "bSize", "bStage", "alphaVegRep","bVegRep")

# MCMC settings
ni <- 10000
nt <- 6
nb <- 5000
nc <- 2

# call JAGS from R
res <- jags(jags.data, inits = NULL, parameters, 
            "centered_pepeSize.jags", n.chains = nc, n.thin = nt, 
            n.iter = ni, n.burnin = nb,
            working.directory = getwd()) 

print(res, digits = 3)

## Check convergence 
# trace plots
traplot(res,c("alpha", "bPlot","bSize", "bStage", "alphaVegRep","bVegRep"))
# posterior distributions
denplot(res,c("alpha", "bPlot","bSize", "bStage", "alphaVegRep","bVegRep"))
```

#######################################################################################################

Old try, trying again  
################################################################################################################################################################################################################  
  simple model of survival (grouping by treatment and control)
  Penstemon penlandii model and data to test
  
```{r,"Old wrong"}
modelpepe <- '	
data {
  for(i in 1:N) {
    y[i] ~ dbern(p[i])
    logit(p[i]) <- alpha + bPlot[ct[i]] ## one for each plot
  }
}

  model{
  fake <- 0
  }
  '


# parameters for simulations 
N = 400 # nb of observations
TC = 4
ct <- as.factor(rep(c("trSp","trFa","coSo","coDe"), each = 100)) # predictor, 4,3,2,1
alpha <- mean(c(1.1,1,-1,-2)) ## because centered, looking for difference from mean
bPlot <- c(1.1,1,-1,-2) # index:  coDe = 1, coSo =2,  trFa =3, trSp=4
# parameters are treated as data for the simulation step
# dat <- unlist(lapply(inv.logit(alpha + bPlot[ct]), function(x) rbinom(n = 1, size = 1, prob = x)))

## Or simulated with model
data<-list(N=N,ct=ct, alpha=alpha,bPlot=bPlot)

# Run JAGS but monitor the response variable instead of the parameters because we're checking if the model works 
out <- run.jags(modelpepe, data = data,monitor=c("y"),sample=1, n.chains=1, summarise=FALSE)

Simulated <- coda::as.mcmc(out)
## spit out y for each data point
dat <- as.vector(Simulated)

```

 <https://mmeredith.net/blog/2017/Categories_in_JAGS.htm>     
# fit model center caegorical coefficients (subtract out the mean)
```{r}
# specify model in BUGS language
modelpepe <- 	
paste("	
model {
  for(i in 1:N) {
    y[i] ~ dbern(p[i])
    logit(p[i]) <- alpha + bPlot[ct[i]]
    }

  # Priors
  alpha ~ dunif(-5,5) ## because it is centered 
    for(i in 1:4) {
      Plottemp[i] ~ dnorm(0, 0.1)   ## find the mean across plotTypes
    }
  bPlot <- Plottemp - mean(Plottemp) ## resulting in change in centered Plot for each type
}")
writeLines(modelpepe,"centered_pepe.jags")

# data, dat from simulation
jags.data <- list(y = dat, ct = ct, N = length(dat))

# initial values
inits <- function(){list(alpha = rbeta(1,1,1), beta = rnorm(1))}  

# parameters monitored
parameters <- c("bPlot", "alpha")

# MCMC settings
ni <- 10000
nt <- 6
nb <- 5000
nc <- 2

# call JAGS from R
res <- jags(jags.data, inits = NULL, parameters, 
            "centered_pepe.jags", n.chains = nc, n.thin = nt, 
            n.iter = ni, n.burnin = nb,
            working.directory = getwd()) 

print(res, digits = 3)

## Check convergence 
# trace plots
traplot(res,c("alpha", "bPlot"))
# posterior distributions
denplot(res,c("alpha", "bPlot"))
```

#######################################################################################################
